### 2013-09-30 src2/bb/paper.tex

%-----------------------------------------------------------------------------
%
%               Template for sigplanconf LaTeX Class
%
% Name:         sigplanconf-template.tex
%
% Purpose:      A template for sigplanconf.cls, which is a LaTeX 2e class
%               file for SIGPLAN conference proceedings.
%
% Guide:        Refer to "Author's Guide to the ACM SIGPLAN Class,"
%               sigplanconf-guide.pdf
%
% Author:       Paul C. Anagnostopoulos
%               Windfall Software
%               978 371-2316
%               paul@windfall.com
%
% Created:      15 February 2005
%
%-----------------------------------------------------------------------------

\documentclass{sigplanconf}

% The following \documentclass options may be useful:
%
% 10pt          To set in 10-point type instead of 9-point.
% 11pt          To set in 11-point type instead of 9-point.
% authoryear    To obtain author/year citation style instead of numeric.

\usepackage{amsmath}

%added JS
\usepackage{graphicx}
\usepackage{placeins}
\usepackage{url}
\usepackage{xcolor}
\usepackage{dialogue}
\usepackage{fancyvrb}
\usepackage[T1]{fontenc}

\definecolor{light-gray}{gray}{0.85}
\newcommand{\docname}{Bj\"{o}rn}
\newcommand{\studname}{Benny}

\begin{document}

\exclusivelicense
\conferenceinfo{ICFP~'13}{September 25--27,2013, Boston, MA, USA} 
\copyrightyear{2013} 
\copyrightdata{978-1-4503-2326-0/13/09} 
\doi{2500365.2500611}

\titlebanner{banner above paper title}        % These are ignored unless
\preprintfooter{short description of paper}   % 'preprint' option specified.

%\title{Reifiable and Compositional Monadic Embedded Languages}
\title{Simple and Compositional Reification of Monadic Embedded Languages}
\subtitle{Functional Pearl}

\authorinfo{Josef Svenningsson \and Bo Joel Svensson}
           {Chalmers University of Technology}
           {\{josefs,joels\}@chalmers.se}

\maketitle

\begin{abstract}
When writing embedded domain specific languages in Haskell, it is
often convenient to be able to make an instance of the {\tt
Monad} class to take advantage of the do-notation and the extensive
monad libraries. Commonly it is desirable to compile such languages 
rather than just interpret them.
This introduces the problem of monad {\em reification}, i.e. observing 
the structure of the monadic computation.
We present a solution to the monad reification problem and
illustrate it with a small robot control language.
Monad reification is not new but the novelty
of our approach is in its directness, simplicity and compositionality.
\end{abstract}

%\category{CR-number}{subcategory}{third-level}
\category{D.3.2}{Programming Languages}{Language Classifications}[Applicative (functional) languages]
\category{D.3.4}{Programming Languages}{Processors}[Code generation]

\keywords
Compiling, embedded language, reification, monads

%\section{Plan}
%
%\begin{itemize} 
%\item The language: Robots
%\item Problem statement 
%  \begin{itemize}
%  \item Want to use DO-notation
%  \item Examples
%        Spirals

%  \item Want to be able to reify a syntax tree
%  \item why ? (Compile and transfer code to actual physical robot) 
%  \end{itemize}

%\item What is the most Naive way to approach this? 
%  \begin{itemize}
%  \item Add return and bind constructors + monad instance. 
%  \end{itemize}
%\item How can this Possibly work? 
%  \begin{itemize} 
%  \item Evaluation 
%  \item Compilations
%  \end{itemize}
%\item Compositional! 
%  \begin{itemize} 
%  \item Data types a'la carte! : NO 
%  \item Will Use the CompData approach. (which is based on a'la carte though) 
%  \end{itemize}
%\item Monad Laws 
%  \begin{itemize} 
%  \item eval (return a \verb!>>=! f) = eval (f a) 
%  \end{itemize}
%\item Related (if at all) 
%
%\end{itemize}

\input{body}
%\input{partII}
%\input{partIII}

%\appendix
%\section{Appendix Title}

%This is the text of the appendix, if you need one.

\acks
We would like to thank Emil Axelsson for very valuable help 
during the implementation of the methods used in this pearl. We 
thank Mary Sheeran for suggesting the names \docname~ \& \studname. 
The ICFP reviewers provided many helpful suggestion which has improved the 
paper.

This research has been funded by the Swedish Foundation for
Strategic Research (which funds the Resource Aware Functional 
Programming (RAW FP) Project) and by the Swedish Research Council.

%Acknowledgments, if needed.

% We recommend abbrvnat bibliography style.

\bibliographystyle{abbrvnat}

% The bibliography should be embedded for final submission.
\begin{thebibliography}{9}
\providecommand{\natexlab}[1]{#1}
\providecommand{\url}[1]{\texttt{#1}}
\expandafter\ifx\csname urlstyle\endcsname\relax
  \providecommand{\doi}[1]{doi: #1}\else
  \providecommand{\doi}{doi: \begingroup \urlstyle{rm}\Url}\fi

\bibitem[Axelsson()]{syntactic}
E.~Axelsson.
\newblock Syntactic.
\newblock http://hackage.haskell.org/package/syntactic.

\bibitem[Axelsson(2012)]{axelsson2012generic}
E.~Axelsson.
\newblock A generic abstract syntax model for embedded languages.
\newblock In \emph{Proceedings of the 17th ACM SIGPLAN international conference
  on Functional programming}, pages 323--334. ACM, 2012.

\bibitem[Bahr and Hvitved(2012)]{CompData}
P.~Bahr and T.~Hvitved.
\newblock Parametric compositional data types.
\newblock In \emph{MSFP}, pages 3--24, 2012.

\bibitem[Claessen et~al.(2012)Claessen, Sheeran, and Svensson]{Obsidian}
K.~Claessen, M.~Sheeran, and B.~J. Svensson.
\newblock Expressive {A}rray {C}onstructs in an {E}mbedded {GPU} {K}ernel
  {P}rogramming {L}anguage.
\newblock In \emph{Proceedings of the 7th workshop on Declarative aspects and
  applications of multicore programming}, DAMP '12, pages 21--30, New York, NY,
  USA, 2012. ACM.
\newblock ISBN 978-1-4503-1117-5.

\bibitem[Farmer and Gill(2012)]{Farmer}
A.~Farmer and A.~Gill.
\newblock {Haskell} {DSL}s for interactive web services.
\newblock In \emph{1st International Workshop on Cross-model Language Design
  and Implementation}, Sep 2012.
\newblock (published on workshop website,
  http://workshops.inf.ed.ac.uk/xldi2012/).

\bibitem[Hudak(1999)]{HUDAK}
P.~Hudak.
\newblock \emph{The Haskell School of Expression: Learning Functional
  Programming through Multimedia}.
\newblock Cambridge University Press, New York, NY, USA, 1999.
\newblock ISBN 0521643384.

\bibitem[Persson et~al.(2012)Persson, Axelsson, and Svenningsson]{Generic}
A.~Persson, E.~Axelsson, and J.~Svenningsson.
\newblock Generic monadic constructs for embedded languages.
\newblock In \emph{Proceedings of the 23rd international conference on
  Implementation and Application of Functional Languages}, IFL'11, pages
  85--99, Berlin, Heidelberg, 2012. Springer-Verlag.

\bibitem[Sculthorpe et~al.(2013)Sculthorpe, Bracker, Giorgidze, and
  Gill]{sculthorpe2013constrained}
N.~Sculthorpe, J.~Bracker, G.~Giorgidze, and A.~Gill.
\newblock The constrained-monad problem.
\newblock In \emph{Proceedings of the 18th ACM SIGPLAN international conference
  on Functional programming}. ACM, 2013.

\bibitem[Swierstra(2008)]{Wouter}
W.~Swierstra.
\newblock Data types \`{a} la carte.
\newblock \emph{Journal of Functional Programming}, 18\penalty0 (4):\penalty0
  423, 2008.

\end{thebibliography}

%\bibliography{paper}

%\begin{thebibliography}{}
%\softraggedright

%\bibitem[Smith et~al.(2009)Smith, Jones]{smith02}
%P. Q. Smith, and X. Y. Jones. ...reference text...

%\end{thebibliography}

\end{document}