### 2013-09-30 src2/bb/partII.tex

\section{Compositional}

One of the advantages of the monad reification technique we have
presented is that it is compositional. The monadic constructs can be
dealt with separately from all the primitives of the embedded
langauges.

A consequence of the compositionality is that we can use techniques
such as compdata \cite{} to decompose the definition of DSLs into
smaller reusable building blocks. Figure \ref{fig:compdata} shows a
compositional implementation of the monadic constructs, separately
from any other constructs. Firstly, our type \verb!MonadT! define

The type \verb!Cxt! is defined in the
package compdata and ties the recursive knot.

\begin{figure}
\begin{verbatim}
data MonadT r a where
  Return :: a -> MonadT r a
  Bind   :: r a -> (a -> r b) -> MonadT r b

instance (MonadT :<: f) => Monad (Cxt h f a) where
  return a = iReturn a
  m >>= f  = iBind m f
\end{verbatim}
\label{fig:compdata}
\caption{Compositional definition monads}
\end{figure}