### 2013-09-30 src2/bb/partIII.tex

\section{Limitations}

%\begin{minipage}[]{\linewidth}
\begin{dialogue}
\speak{\docname{}} There is still one problem with your method though.

\speak{\studname{}} What's that?

\speak{\docname{}} When we make instances of the Monad class we expect
certain laws to hold. \direct{\docname{} writes the laws on his whiteboard}

\end{dialogue}
%\end{minipage}

\begin{small}
\begin{verbatim}
m >>= return    = m
return a >>= f  = f a
(m >>= f) >>= g = m >>= \a -> f a >>= g
\end{verbatim}
\end{small}

%\begin{minipage}[]{\linewidth}
\begin{dialogue}
\speak{\docname{}} These laws clearly don't hold for your instance. Take the
first law for instance. The left hand side will have extra \texttt{Bind} and
\texttt{Return} constructors compared to the right hand side.

\speak{\studname{}} Hmmm. I hadn't really thought about that. But adding an
extra return at the end of a computation shouldn't make a difference
in my implementation.

\speak{\docname{}} So, you're saying that your implementation actually obeys
the first monad law?

\speak{\studname{}} Well, at least when I run my programs I will never see
any difference between the left hand side and the right hand side.

\speak{\docname{}} Aha, so what you're saying is that if we compare the
semantics of programs rather than comparing the programs themselves
then we get some useful laws. \direct{\docname{} scribbles some new laws
  on the whiteboard}

\end{dialogue}
%\end{minipage}

\begin{small}
\begin{verbatim}
eval (m >>= return)    = eval m
eval (return a >>= f)  = eval (f a)
eval ((m >>= f) >>= g) = eval (m >>= \a -> f a >>= g)
\end{verbatim}
\end{small}

%\begin{minipage}[]{\linewidth}
\begin{dialogue}

\speak{\docname{}} These laws a morally the same as the monad
laws, especially if we don't let the user of the robot language ever
compare terms in the language.

\speak{\studname{}} Yes, that captures my intuition very well.

\speak{\docname{}} Ok, good. Can you prove these equations?
\end{dialogue}
%\end{minipage}

\emph{Prove the monad laws}\newline \newline

% \docname{} calls upon \studname{} to tally up the limitations of his approach

% Monad Laws (Do the hold ?) 
% Infinite programs 
% Only certains ``kinds'' of reification!

\section{Benefits}

\section{Epilogue}

The trick which we've shown in this paper scales up nicely to larger
embedded languages. The language Obsidian already use it.

% we should always point of the positives as well!

% we should always point of the positives as well!