### 2013-09-30 src2/expressive/abstract.tex

%\begin{abstract}
%This is the text of the abstract.

\section*{Abstract}
Graphics Processing Units (GPUs) are powerful computing devices 
that with the advent of CUDA/OpenCL are becomming useful for general 
purpose computations. 
Obsidian is an embedded domain specific language that generates CUDA kernels
from functional descriptions.
A symbolic array construction allows us to guarantee that intermediate
arrays are fused away. However, the current array construction has
some drawbacks; in particular, arrays cannot be combined efficiently.
We add a new type of {\em push arrays} to the existing Obsidian system in 
order to solve this problem. The two array types complement each other,
and enable the definition of combinators that both
take apart and combine arrays, and that result in efficient generated code.
This extension to Obsidian is demonstrated on a sequence of
sorting kernels, with good results.
The case study also illustrates the use of combinators for expressing
the structure of parallel algorithms.
The work presented is preliminary, and 
the combinators presented must be generalised.
However, the raw speed of the generated kernels bodes well.
%\end{abstract}