### 2013-09-30 src2/expressive/discussion.tex

\section {Discussion}

%% ** TODO. Mention CSE. warp-related opt.

Push arrays form a new approach to array representation in DSELs.
We do not know of similar approaches in the literature, despite
the fact that the notions of demand and data flow
may feel familiar to the reader who considers Pull and Push arrays.
The addition of Push arrays to Obsidian seems highly beneficial. With 
this new feature, the user gains finer grained control over the code generated
and the resulting CUDA kernels perform considerably better than before.
This 
was illustrated in the series of sorters explored in section~\ref{sec:MARY}.
The performance of {\tt vsort} is sufficiently good that it
can be used as a first phase in a larger sorter (written in CUDA) that
can sort 16M elements in 96 ms, while an i7-920 CPU takes
around 2740 ms. Further speed improvements look possible, both in the coordination code
and in the kernels. An obvious next step would be to investigate the generation of the {\tt iSwap} and
{\tt vSwap} kernels from Obsidian. (This is not currently possible because of assumptions that we made about the interfaces to kernels and about how {\em thread ids} are used. We will look into ways to relax
our assumptions.)

The series of kernels also illustrates how the use of combinators brings a form of reuse, and makes design exploration easier. Our experience of using similar combinators in the Lava hardware
description language~\cite{LavaSorter} was that a relatively small set of combinators went a long way. So, although we introduced three
combinators here, {\tt ilv}, {\tt vee} and {\tt ilvVee}, which
includes the other two, we do not believe that every new kernel development exercise would demand a completely new set of combinators.
We expect to provide the user with a well-documented set of combinators,
so that users can get access to this style of programming without having
to develop their own combinators, and without having to think too much
about bit-hacking.
The bit-manipulation approach chosen to define our combinators automatically
created functions that apply to sub-sequences of the input that
are of an appropriate length.

In this paper, we made combinators for the special case of two input, two
output operations (built from two two-input funtions that we typically called
{\tt f} and {\tt g}). 
This approach should be generalised to deal with blocks that have $2^k$ inputs
and outputs.
Also, we made a compound combinator from {\tt ilv} and {\tt vee},
but generalising to more than two input components would allow
for composing combinators, and indeed for recursive descriptions
that could be unrolled. Then, ignoring {\tt syncs}, a recursive description of
{\tt vsort} could be something like 
\begin{codesize}
\begin{verbatim}
vsortR 0 = id
vsortR n = bpmergeR n . ilv2 1 (vsortR (n-1))
\end{verbatim}
\end{codesize}
\noindent
It would then be necessary to optimise the code generated from
multiple applications of {\tt ilv2 1}, for example, whereas here
we have forced the user to figure out both the unrolling and the combinations.
Moving to more general combinators would also give the opportunity to
provide predefined combinators that capture more of the commonly used threading patterns (for instance $k$ indices per thread rather than the $1$ and $2$ shown here).

The integration of Push arrays into Obsidian raises some new questions.
Previously, there was a direct correspondence between the length 
of an array and the number of threads used to compute it, which
allowed the user to write an initial program without worrying about
threads at all, and then to tweak the Obsidian program if he was not satisfied
with the threading behaviour of the resulting kernel. 
Now, as can be seen in the {\tt catArrayPs} example and in the sorters, this correspondence
can be broken using Push arrays. The {\tt catArrayPs} example and two
of the sorters use half as many threads as the 
number of elements. For users who are very concerned about the speed of
the generated kernels, getting this control through using Push arrays in
a particular pattern is clearly a good thing. But adding a second, different
way to control thread use in the generated code certainly complicates matters, and further case studies are needed to confirm that the complication pays off.

The addition of Push arrays also adds the possibility to include potentially 
unsafe operations in Obsidian, for example by writing multiple array elements to the same index, or by discarding elements.
This new expressiveness will have to be carefully controlled. On the positive side, it offers the possibility to encode functions like {\tt filter} from Haskell
that
are simply not expressible using only Pull arrays. Being able to implement {\tt filter} would make programming kernels in obsidian feel much more like programming in Haskell -- a welcome loosening of the strait-jacket.
Once that is done, it will be time to develop a very simple coordination language to allow programming of entire GPU applications that make use of
the kind of small kernel building blocks developed here.