### 2013-09-30 src2/expressive/paperFinal.tex

%-----------------------------------------------------------------------------
%
%               Template for sigplanconf LaTeX Class
%
% Name:         sigplanconf-template.tex
%
% Purpose:      A template for sigplanconf.cls, which is a LaTeX 2e class
%               file for SIGPLAN conference proceedings.
%
% Author:       Paul C. Anagnostopoulos
%               Windfall Software
%               978 371-2316
%               paul@windfall.com
%
% Created:      15 February 2005
%
%-----------------------------------------------------------------------------

\documentclass[]{sigplanconf}

% The following \documentclass options may be useful:
%
% 10pt          To set in 10-point type instead of 9-point.
% 11pt          To set in 11-point type instead of 9-point.
% authoryear    To obtain author/year citation style instead of numeric.

\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{url}

\newcommand{\codesize}{\fontsize{7}{8}\selectfont}

\begin{document}
\conferenceinfo{DAMP'12,} {January 28, 2012, Philadelphia, PA, USA.}
\CopyrightYear{2012}
\copyrightdata{978-1-4503-1117-5/12/01}

%\conferenceinfo{WXYZ '05}{date, City.} 
%\copyrightyear{2005} 
%\copyrightdata{[to be supplied]}

\titlebanner{}        % These are ignored unless
\preprintfooter{}   % 'preprint' option specified.

%\title{Towards an Implementation of Push Arrays in Obsidian}
\title{Expressive Array Constructs in an Embedded GPU Kernel Programming Language}
%\subtitle{Subtitle Text, if any}

%\authorinfo{Name1}
%           {Affiliation1}
%           {Email1}
\authorinfo{Koen Claessen \and Mary Sheeran \and Bo Joel Svensson}
           {Chalmers University of Technology, 
             Department of Computer Science and Engineering,
             Gothenburg, Sweden}
           {koen@chalmers.se/ms@chalmers.se/joels@chalmers.se}

\maketitle

% ABSTRACT !!! -----------------------------------------------------------------
\begin{abstract}
%This is the text of the abstract.

Graphics Processing Units (GPUs) are powerful computing devices 
that with the advent of CUDA/OpenCL are becomming useful for general 
purpose computations. 
Obsidian is an embedded domain specific language that generates CUDA kernels
from functional descriptions.
A symbolic array construction allows us to guarantee that intermediate
arrays are fused away. However, the current array construction has
some drawbacks; in particular, arrays cannot be combined efficiently.
We add a new type of {\em push arrays} to the existing Obsidian system in 
order to solve this problem. The two array types complement each other,
and enable the definition of combinators that both
take apart and combine arrays, and that result in efficient generated code.
This extension to Obsidian is demonstrated on a sequence of
sorting kernels, with good results.
The case study also illustrates the use of combinators for expressing
the structure of parallel algorithms.
The work presented is preliminary, and 
the combinators presented must be generalised.
However, the raw speed of the generated kernels bodes well.
\end{abstract}

%\category{CR-number}{subcategory}{third-level}
\category{D.3.2}{Programming Languages}{Language Classifications}[Applicative (functional) languages; Concurrent, distributed, and parallel languages]
\category{D.3.4}{Programming Languages}{Processors}[Code generation]

\terms
Languages, Algorithms, Performance

\keywords
Arrays, Data parallelism, Embedded Domain Specific Language, General Purpose GPU programming, Haskell

%\input{intro}
\section{Introduction}

%\input{table}

%----------------------------------------------------------------------------

Graphics Processing Units (GPUs) are parallel computers with hundreds 
to thousands of processing elements. The CUDA and OpenCL languages  
make available the power of the GPU to programmers interested in general 
purpose computations. In CUDA and OpenCL, the programmer writes {\em kernels}, 
Single Program Multiple Data (SPMD) programs that are executed by groups 
of threads on the available processing elements of the GPU.

CUDA and OpenCL are general purpose programming languages, mirroring the 
increased capabilities of a modern GPU to target that domain. However, these 
languages lack compositionality. 
Also, being based in C/C++ means that the core idea in a program may
not be easily visible.

\subsection{Embedded DSLs for GPGPU programming} 
We are aiming for a GPU programming language that is more
concise than mainstream
languages such as CUDA and OpenCL. 
Obsidian is a
domain specific embedded language (DSEL) implemented in Haskell.
When an Obsidian program is run, a representation of the program is created 
as a syntax tree. For more information on EDSL implementation see~\cite{COMPILEEDSL}.
The program representation generated when running an Obsidian program 
is compiled into CUDA code. We are also working on an OpenCL backend.

Our approach is different from that of other Haskell DSELs targetting
GPUs~\cite{ACCELERATE,NIKOLA,BARRACUDA}. 
We do not try to abstract away from 
all the peculiarities of GPU programming, but rather provide a higher 
level language in which to experiment with them.
For instance,
Accelerate provides a standard set of basic operations such as 
{\tt map}, {\tt reduce} and {\tt zipWith} as built in skeletons, implemented
with the help of small, predefined, hand-tuned CUDA kernels~\cite{ACCELERATE}. 
Obsidian, on the other hand, allows the user to experiment with the
{\em generation} of small kernels for fixed size array inputs
from higher level descriptions.
It is intended to allow the user to play with the kinds of tradeoffs that are important
when writing such high performance building blocks; in this paper, the main consideration
is the number of array elements of the input and output that are manipulated
by a single thread in the generated CUDA code.
An important aspect of Obsidian is the symbolic array representation used, along
with its associated {\tt sync} operation. As we shall see, the {\tt sync} operation
allows the programmer to guide code generation and control parallelism and thread
use~\cite{JSLIC}.

%Today there are number of EDSLs written in Haskell targeting 
%GPUs~\cite{ACCELERATE,NIKOLA,BARRACUDA}. We justify our work on yet 
%another by having a slightly different approach. We do not try to abstract away from 
%all the peculiarities of GPU programming but rather giving a higher 
%level language in which to experiment with them.
%trying to appeal to a slightly different target audience. 
%We try to cater not for the advanced functional programmer who wants to 
%run his programs on GPUs, but rather to offer a higher level 
%language to the CUDA/OpenCL hacker. 
%In this language we want the user to be 
%able to quickly experiment with the kind of tradeoffs that are important 
%when writing a high performance GPU kernel.

%Unlike the other approaches that are complementary to our Haskell GPU EDSL, 
%Obsidian does not provide the usual set of basic operations such as 
%{\tt Map}, {\tt Reduce} and {\tt ZipWith} as built in skeletons. We 
%want a language where the programmer can experiment with implementing 
%his or her own versions of these basic operations. So far, we make
%this possible in Obsidian through a {\tt Sync} operation and the Array 
%representation we use. The {\tt Sync} operation, using target domain 
%nomenclature, is the programmers tool to guide code generation and 
%obtain parallelism. The presentation of Obsidian here will be somewhat 
%cursory, for a more complete view of the various versions and approaches 
%we have tried, see~\cite{JSLIC}.
%More details on this later.

In Obsidian, a kernel that sums an array can be expressed as: 
\begin{codesize}
\begin{verbatim}
sum :: Array IntE -> Kernel (Array IntE) 
sum arr | len arr == 1 = return arr
        | otherwise    = (pure (fmap (uncurry (+)) . pair) 
                          ->- sync 
                          ->- sum) arr                       
\end{verbatim}
\end{codesize}
The result of running this kernel on an eight element input array, 
{\tt runKernel sum (namedArray ``input'' 8)}, is an intermediate representation 
of the computation (shown in slightly pretty-printed form): 
\begin{codesize}
\begin{verbatim}
arr0 = malloc(16)
par i 4 {
arr0[i] = ( +  input[( *  i 2 )] input[( +  ( *  i 2 ) 1 )] );
}Sync
arr1 = malloc(8)
par i 2 {
arr1[i] = ( +  arr0[( *  i 2 )] arr0[( +  ( *  i 2 ) 1 )] );
}Sync
arr2 = malloc(4)
par i 1 {
arr2[i] = ( +  arr1[( *  i 2 )] arr1[( +  ( *  i 2 ) 1 )] );
}Sync
\end{verbatim}
\end{codesize}
The named intermediate arrays in this representation are then laid 
out in GPU shared memory and CUDA code can be generated (here for arrays of length eight)\footnote{An alignment qualifier for shared memory 
has been omitted to save space in the listings showing generated code}: 
%\pagebreak
\begin{codesize}
\begin{verbatim}
__global__ void sum(int *input0,int *result0){
  unsigned int tid = threadIdx.x;
  unsigned int bid = blockIdx.x;
  extern __shared__ unsigned char sbase[];
  (( int *)sbase)[tid] = 
    (input0[((bid*8)+(tid*2))]+
     input0[((bid*8)+((tid*2)+1))]);
  __syncthreads();
  if (tid<2){
    (( int *)(sbase + 16))[tid] = 
      ((( int *)sbase)[(tid*2)]+
       (( int *)sbase)[((tid*2)+1)]);
  }
  __syncthreads();
  if (tid<1){
    (( int *)sbase)[tid] = 
      ((( int *)(sbase+16))[(tid*2)]+
       (( int *)(sbase+16))[((tid*2)+1)]);
  }
  __syncthreads();
  if (tid<1){
    result0[(bid+tid)] = (( int *)sbase)[tid];
  }
}
\end{verbatim}
\end{codesize}
%The CUDA code above is generated by issuing the command:
%\begin{codesize}
%\begin{verbatim}
%genKernel ``sum'' sum (namedArray undefined 8 :: Array IntE)
%\end{verbatim}
%\end{codesize}
%\noindent
%and the kernel code is generated for arrays of length eight.

%----------------------------------------------------------------------------
\subsection{Arrays in Obsidian}

An array is represented by an indexing function and a length:
\begin{codesize} 
\begin{verbatim}
data Array a = Array (UWordE -> a) Word32 
\end{verbatim}
\end{codesize}
This array representation has served us well. It has these properties:
\begin{itemize} 
\item Fusion of operations is automatic.
\item It naturally describes a data-parallel computation suitable for CUDA/OpenCL generation.
\item Many basic operations can be implemented: {\tt map}, {\tt zipWith} etc.
\end{itemize}
Using this array representation in a DSEL is not new; the first occurence that we know of is in
Pan~\cite{PAN}. Similar array representations have also later been 
used in Feldspar~\cite{FELDSPAR}, and more recently also in the Repa library~\cite{REPA}.
Functions for indexing and getting the length of arrays are as follows:
\begin{codesize} 
\begin{verbatim}
(!) :: Array a -> UWordE -> a 
(Array ixf _) ! ix = ixf ix

len :: Array a -> Word32
len (Array _ n) = n 
\end{verbatim}
\end{codesize}
%Implementing a {\em Functor} instance for the {\tt Array} datatype is possible:
A {\em Functor} instance for the {\tt Array} datatype is
\begin{codesize} 
\begin{verbatim}
instance Functor Array where 
  fmap f arr = Array (\ix -> f (arr ! ix)) (len arr) 
\end{verbatim}
\end{codesize}
Now, composed applications of {\tt fmap} will be automatically fused. This is 
illustrated in the example program below and the CUDA generated
from it.  
\begin{codesize} 
\begin{verbatim}
mapFusion :: Array IntE -> Kernel (Array IntE) 
mapFusion = pure (fmap (+1) . fmap (*2)) 
\end{verbatim}
\end{codesize}

\begin{codesize} 
\begin{verbatim}
__global__ void mapFusion(int *input0,int *result0){
  unsigned int tid = threadIdx.x;
  unsigned int bid = blockIdx.x;

  result0[((bid*32)+tid)] = ((input0[((bid*32)+tid)]*2)+1);
}
\end{verbatim}
\end{codesize}
Both of these code listings need explanation. In the Haskell 
code, {\tt mapFusion} has type {\tt Array IntE -> Kernel (Array IntE)};
{\tt Kernel} is a state monad that accumulates CUDA code as well as provides 
new names for intermediate arrays. Neither of these features of the monad 
is activated by this example though. The {\tt pure} function is defined 
using the monad's {\tt return} as {\tt pure f a = return (f a)}. In this case, 
it lifts a function of type {\tt Array IntE -> Array IntE} into 
a kernel.

The generated CUDA code computes the result array using a number of threads 
equal to the length of that array. In this case, the kernel was generated
to deal with arrays of length $32$. The important detail to notice in the 
CUDA code is that there is no intermediate array created between the 
{\tt (*2)} and the {\tt (+1)} operations.

The {\tt mapFusion} example could just as well have been implemented using 
the kernel sequential composition combinator, {\tt ->-}. 
\begin{codesize} 
\begin{verbatim}
mapFusion :: Array IntE -> Kernel (Array IntE) 
mapFusion = pure (fmap (*2)) ->- pure (fmap (+1)) 
\end{verbatim}
\end{codesize}
Exactly the same CUDA code is then generated.

In some cases, it is necessary to force computation of 
intermediate arrays. This can be used to share partial computations between 
threads and to expose parallelism. In Obsidian, the tool for this
is called {\tt sync}, a built-in kernel. Using {\tt sync} as follows prevents fusion of the
two operations:
\begin{codesize} 
\begin{verbatim}
mapUnFused :: Array IntE -> Kernel (Array IntE) 
mapUnFused = pure (fmap (*2)) ->- sync ->- pure (fmap (+1))
\end{verbatim}
\end{codesize}
The generated CUDA code now stores an intermediate result in local shared
memory before moving on. 
\begin{codesize} 
\begin{verbatim}
__global__ void mapUnFused(int *input0,int *result0){
  unsigned int tid = threadIdx.x;
  unsigned int bid = blockIdx.x;
  extern __shared__ unsigned char sbase[];
  (( int *)sbase)[tid] = (input0[((bid*32)+tid)]*2);
  __syncthreads();
  result0[((bid*32)+tid)] = ((( int *)sbase)[tid]+1);
}
\end{verbatim}
\end{codesize}
Intermediate arrays are laid out in the {\tt sbase} array in shared memory. 
Since we may store arrays of many different types in the same locations 
of the shared memory at different times during the execution, the type casts
used in the code above are necessary.

%----------------------------------------------------------------------------
\subsection{Sync and parallelism}

The {\tt sync} operation also enables the writing of parallel reduction 
kernels. A reduction operation is an operation that takes an array as input
and
produces a singleton array as output.

First, we define {\tt zipWith} and {\tt halve} on Obsidian arrays.
\begin{codesize} 
\begin{verbatim}
zipWith :: (a -> b -> c) -> Array a -> Array b -> Array c
zipWith op a1 a2 = Array (\ix -> (a1 ! ix) `op` (a2 ! ix)) 
                   (min (len a1) (len a2))

splitAt :: Word32 -> Array a -> (Array a, Array a) 
splitAt n arr = 
  (Array (\ix -> arr ! ix) n , 
   Array (\ix -> arr ! (ix + fromIntegral n)) (len arr - n))

halve arr = splitAt ((len arr) `div` 2) arr
\end{verbatim}
\end{codesize}
A reduction kernel that takes an array whose length is a power of two 
and gives an array of length one can be defined recursively. 
Defining kernels recursively results in completely unrolled CUDA kernels,
and kernel input size must be known at compile time.
The approach to reduction taken here is to split the input array 
into two halves and then apply {\tt zipWith} of the combining function to the 
two halves, repeating the process until the length is one. 
\begin{codesize} 
\begin{verbatim}
reduceS :: (a -> a -> a) -> Array a -> Kernel (Array a) 
reduceS op arr | len arr == 1 = return arr
               | otherwise    = 
                 (pure ((uncurry (zipWith op)) . halve)
                  ->- reduceS op) arr
\end{verbatim}
\end{codesize}
%This Obsidian Kernel takes an array of length $2n$ as input and the output 
Since the output of this kernel is of length one, and the number 
of elements in the output array specifies the number of threads used to compute it, 
this function, {\tt reduceS}, defines a sequential reduction. 
The generated code for arrays of length eight is 
\pagebreak
%The code is generated with the 
%command {\tt genKernel ``reduceSAdd'' (reduce (+)) (namedArray undefined 8 :: Array IntE)}, 
%meaning that a CUDA reduce-with-addition kernel for arrays of length eight is generated. 
\begin{codesize} 
\begin{verbatim}
__global__ void reduceSAdd(int *input0,int *result0){
  unsigned int tid = threadIdx.x;
  unsigned int bid = blockIdx.x;

  result0[(bid+tid)] = 
    (((input0[((bid*8)+tid)]+
       input0[((bid*8)+(tid+4))])+
      (input0[((bid*8)+(tid+2))]+
       input0[((bid*8)+((tid+2)+4))]))+
     ((input0[((bid*8)+(tid+1))]+
       input0[((bid*8)+((tid+1)+4))])+
      (input0[((bid*8)+((tid+1)+2))]+
       input0[((bid*8)+(((tid+1)+2)+4))])));
}
\end{verbatim}
\end{codesize}
Sequential reduction is not very interesting for GPU execution, but 
the fix is simple. A well placed use of {\tt sync} indicates that 
we want to compute, after each {\tt zipWith} phase, the intermediate 
arrays using as many threads as that intermediate array is long. 
The effect is shown in the code below. 
\begin{codesize} 
\begin{verbatim}
reduce :: Syncable Array a 
          => (a -> a -> a) -> Array a -> Kernel (Array a)
reduce op arr | len arr == 1 = return arr
              | otherwise    = 
                (pure ((uncurry (zipWith op)) . halve)
                 ->- sync 
                 ->- reduce op) arr
\end{verbatim}
\end{codesize}
The CUDA code for reduction with addition on eight
elements is
\begin{codesize} 
\begin{verbatim}
__global__ void reduceAdd(int *input0,int *result0){
  unsigned int tid = threadIdx.x;
  unsigned int bid = blockIdx.x;
  extern __shared__ unsigned char sbase[];
  (( int *)sbase)[tid] = 
    (input0[((bid*8)+tid)]+
     input0[((bid*8)+(tid+4))]);
  __syncthreads();
  if (tid<2){
    (( int *)(sbase + 16))[tid] = 
      ((( int *)sbase)[tid]+
       (( int *)sbase)[(tid+2)]);   
  }
  __syncthreads();
  if (tid<1){
    (( int *)sbase)[tid] = 
      ((( int *)(sbase+16))[tid]+
       (( int *)(sbase+16))[(tid+1)]);    
  }
  __syncthreads();
  if (tid<1){
    result0[(bid+tid)] = (( int *)sbase)[tid];  
  }
}
\end{verbatim}
\end{codesize}           
In this generated CUDA, three phases can be identified. 
The first uses four threads to compute a four element intermediate array;
the second uses two threads, and so on. At the very end, a single thread
copies the result from local shared memory to global memory.

\subsection{Drawbacks of Obsidian Arrays}
\label{sec:Drawbacks}

%*** THE STUFF BELOW IS IN CONFLICT WITH WHAT I ADDED ABOVE!
%When using Obsidian we have available to us a toolbox of higher order
%functions that are familiar to functional programmers. One example of such a 
%function is {\tt zipWith}.

%\begin{codesize} 
%\begin{verbatim}
%example1 :: (Array IntE, Array IntE) 
%            -> Kernel (Array IntE)
%example1 (arr1,arr2) = return$ zipWith (+) arr1 arr2
%\end{verbatim}
%\end{codesize}

%The body of this function looks very familiar to Haskell programmers. The 
%type of the function however is a little bit different than what we are 
%used to from everyday Haskell usage. In Obsidian we have a concept called 
%a Kernel. Computations of types {\tt a -> Kernel b} can be turned into 
%runnable GPU code by providing an input. 
%*** VERY VAGUE AND ODD 
%*** TODO: Explain kernel ?

%The following code listing shows the CUDA code we can generate from 
%the Obsidian program given above:

% *** TODO: Needs a code size between tiny and small.

%\begin{codesize}
%\begin{verbatim}
%__global__ void example1(int *input0,int *input1,int *result0){
%  unsigned int tid = threadIdx.x;
%  unsigned int bid = blockIdx.x;

%  result0[tid] = (input0[((bid*32)+tid)]+
%                  input1[((bid*32)+tid)]);

%}
%\end{verbatim}
%\end{codesize}

%The CUDA code above writes into a result array the pairwise sums from 
%its two input arrays, as expected. 
%*** There are many details here. tid, bid, 32. explain now or push to later?

% \subsection{Obsidian arrays}

%In Obsidian an array is represented by an indexing function and a length. 
%\begin{codesize}
%\begin{verbatim}
%data Array a = Array (IxExp -> a) Word32
%\end{verbatim}
%\end{codesize}

%And {\tt IxExp} is defined as {\tt UWordE}, that is an expression 
%representing a unsigned 32bit value. The length of an array is represented
%with a unsigned 32bit value as well but here it has been beneficial 
%to use a built in Haskell type in order to be able to use this 
%information at program generation time rather than program run time.

%This array representation specifies how to compute each index of an Arrays. 
%If we are interested in the value at location 5 in an array all we need 
%to do is apply the indexing function to the literal expression 5 and 
%we get a description of how to compute that value.

%This array representation have been working fairly well in many situations. 
%In {\tt example1} where we used {\tt zipWith} we got just about the resulting
%CUDA code that a CUDA programmer would write himself. For some other examples 
%the situation is much worse.

The previous subsection described positive aspects of the array 
representation that we have used so far. There are, however, 
circumstances in which this Array representation is too restricted.

Take the problem of concatenating two arrays. Using the array representation 
described above, the only way to concatenate two arrays is to introduce 
a conditional into the indexing function. If {\tt f} and {\tt g} are the 
indexing functions of two arrays that are to be concatenated, and {\tt n1}
is the length of the first array, the indexing 
function of the result must be
\begin{codesize}
\begin{verbatim}
new ix = if (ix < n1) 
         then f ix 
         else g (ix - n1)
\end{verbatim}
\end{codesize}
The following program
concatenates two arrays: 
\begin{codesize}
\begin{verbatim}
catArrays :: (Array IntE, Array IntE) -> Kernel (Array IntE)
catArrays = pure conc
\end{verbatim}
\end{codesize}
%(arr1,arr2) = return$ conc (arr1,arr2)
When it is used to generate a CUDA kernel that concatenates
two arrays of length $16$, the following code is the result: 
\begin{codesize}
\begin{verbatim}
__global__ void catArrays(int *input0,int *input1,int *result0){
  unsigned int tid = threadIdx.x;
  unsigned int bid = blockIdx.x;

  result0[((bid*32)+tid)] = 
    (tid<16) ? input0[((bid*16)+tid)] : 
               input1[((bid*16)+(tid-16))];
}
\end{verbatim}
\end{codesize}
%__global__ void catArrays(int *input0,
%                          int *input1,
%                          int *result0){
%  unsigned int tid = threadIdx.x;
%  unsigned int bid = blockIdx.x;
%  
%  result0[((bid*64)+tid)] = 
%    (tid<32) ? input0[((bid*32)+tid)] : 
%               input1[((bid*32)+(tid-32))];
%  
%}
Now, conditionals like these are {\em bad} in code 
to execute on a GPU, with its wide-SIMD data-parallel model. Separating
the operation into two assignments and using half as many threads gives
much higher performance.
\begin{codesize}
\begin{verbatim}
__global__ void catArraysByHand(int *input0,
                                int *input1,
                                int *result0){
  unsigned int tid = threadIdx.x;
  unsigned int bid = blockIdx.x;

  result0[((bid*32)+tid)] = input0[((bid*16)+tid)];
  result0[((bid*32)+tid+16)] = input1[((bid*16)+tid)];  
}
\end{verbatim}
\end{codesize}
There are cases where code with conditionals is not that bad. 
An expert on NVIDIA GPUs in particular may say that code with 
the condition {\tt (tid < 32)} is fine, since 32 is the SIMD width 
of those GPUs. However, any number that is not a multiple of 32 would 
lead to poor performance, so in general this is a problem. 
 Worse still, {\em zipping} two arrays together and then 
{\em unpairing} (to get an array of elements)
leads to code that takes two different paths depending 
on odd or even {\em thread id}. When a GPU executes such code, it 
shuts down half of the threads and computes the two paths in sequence. 
\begin{codesize}
\begin{verbatim}
zippUnpair :: (Array IntE, Array IntE) -> Kernel (Array IntE) 
zippUnpair = pure (unpair . zipp)
\end{verbatim}
\end{codesize}
The {\tt zipp} and {\tt unpair} operations are defined as follows: 
\begin{codesize}
\begin{verbatim}
zipp :: (Array a, Array b) -> Array (a, b)             
zipp (arr1,arr2) = 
     Array (\ix -> (arr1 ! ix, arr2 ! ix)) 
           (min (len arr1) (len arr2))

unpair :: Choice a => Array (a,a) -> Array a
unpair arr = 
    let n = len arr
    in  Array (\ix -> ifThenElse ((mod ix 2) ==* 0) 
                      (fst (arr ! (ix `shiftR` 1)))
                      (snd (arr ! (ix `shiftR` 1)))) (2*n)
\end{verbatim}
\end{codesize}
Code generated from the {\tt zippUnpair} program exhibits really 
poor performance; at any time half of the threads are shut down. 
\begin{codesize}
\begin{verbatim}
__global__ void zippUnpair(int *input0,
                           int *input1,
                           int *result0){
  unsigned int tid = threadIdx.x;
  unsigned int bid = blockIdx.x;

  result0[((bid*64)+tid)] = 
    ((tid%2)==0) ? input0[((bid*32)+(tid>>1))] : 
                   input1[((bid*32)+(tid>>1))];
}
\end{verbatim}
\end{codesize}
If we wrote this CUDA program by hand, we would, again, split it up into 
two phases so that all threads can progress in parallel.

%Even though this is a pretty artificial example, it illustrates the 
%problem well. The situation is just as bad when using the function
%{\tt unpair}. This function takes an array of pairs and returns an 
%array of elements such that position 0 and 1 in the result array holds 
%the values of the pair at index 0 in the input array. Using this operation 
%results in a conditional statement that takes different execution paths 
%on odd and even numbered threads.

The arrays described so far, with an indexing function and a length, have 
been nicknamed {\em Pull arrays} for how they describe how to 
compute an element by {\em pulling} data from a number 
of places. Using just Pull arrays, we have been unable to solve the 
problems described so far in this section. The solution is to add a complementary array type to Obsidian.

%For a more in depth view of the parts of Obsidian that has been showed 
%in this section see~\cite{JSLIC}.

%\input{push} 
\section{Push Arrays}

In order to improve low level control for the programmer, {\em Push
arrays} are added to Obsidian. The old Pull arrays are still available,
along with the new array type.

Some operations, typically involving taking arrays apart,
are easily described using Pull arrays, giving
efficient code. In those cases, using a Push array would add complexity in the 
implementation for no performance benefit.
Other operations cannot be implemented efficiently
with Pull arrays, but Push arrays then provide the solution.
This duality is apparent when looking at operations on Pull arrays such as
{\tt halve} and {\tt conc} (for concatenate). 
The {\tt halve} function is efficient since it introduces no diverging 
conditionals. The {\tt conc} function, on the other hand, introduces conditionals.
Concatening two arrays using the {\tt concP} combinator, implemented on Push arrays, 
allows us to generate the desired code:

\begin{codesize}
\begin{verbatim}
catArrayPs :: (Array IntE, Array IntE) -> Kernel (ArrayP Int)
catArrayPs = pure concP
\end{verbatim}
\end{codesize}

\begin{codesize}
\begin{verbatim}
__global__ void catArrayPs(int *input0,int *input1,int *result0){
  unsigned int tid = threadIdx.x;
  unsigned int bid = blockIdx.x;

  result0[((bid*32)+tid)] = input0[((bid*16)+tid)];
  result0[((bid*32)+(16+tid))] = input1[((bid*16)+tid)];

}
\end{verbatim}
\end{codesize}

\noindent
Compared to the CUDA code for {\tt catArrays}, this kernel 
uses only $16$ threads instead of $32$. At each step of the computation, all 
the threads are fully busy doing exactly the same thing, which is the preferred 
mode of execution on the target platform.

%The performance effects of these unwanted conditionals, compared to code generated 
%using the new Push arrays, will be illustrated in section~\ref{sec:MARY}.

%The {\tt halve} operation has appeared previously and the {\tt pair} 
%operation is its inverse. Implementing the {\tt pair} operation using 
%Pull arrays is very easy and also efficient:

%\begin{codesize}
%\begin{verbatim}
%pair :: Array a -> Array (a,a)
%pair (Array ixf n) = Array (\ix -> (ixf (ix*2),ixf (ix*2+1))) n'
%  where 
%    n' = n `div` 2 
%\end{verbatim}
%\end{codesize}

%It is efficient because it introduces no diverging conditionals. There is
%just a tiny bit of extra arithmetic in the indexing function. On the 
%other hand {\tt unpair}, as we saw, introduced an unnecessary and 
%inefficient conditional. It turns out that {\tt unpair} can be 
%efficiently implemented on a Push array.

\subsection {What are Push Arrays?}

The idea behind Push arrays is to have a way to describe where 
elements are supposed to end up. In some sense, a Push array produces 
a collection of Index/Value pairs. %This makes a Push array a lot more 
%powerful than a Pull array. 
This makes Push arrays complementary to Pull arrays. For example, it 
is possible for a Push array to output several elements at the same index 
(which we probably need to control carefully). Push arrays should permit 
us to provide more expressive operations on arrays to the user, including 
an operation similar to Haskell's {\tt filter} on lists. Here, we consider 
a different advantage of adding push arrays: finer control over patterns of 
thread use in generated code.

%% Will it really give nondeterminism or will the last one written win, or some such?

A Push array consists of three parts: a function in continuation passing style, 
a {\tt Program} datatype and an array datatype.

\begin{codesize}
\begin{verbatim}
type P a = (a -> Program) -> Program 
\end{verbatim}
\end{codesize}

\noindent
For another example of using continuations and a more complete
description of their meaning and application, see~\cite{POORKOEN}.

The {\tt Program} datatype has now been adopted as Obsidian's internal 
representation of CUDA programs.

\begin{codesize}
\begin{verbatim}
data Program
  = Skip
  | forall a. Scalar a => Assign Name UWordE (Exp a) 
  | Par (UWordE -> Program) Word32   
  | Allocate Name Word32 Type 
  | Synchronize 
  | ProgramSeq Program 
               Program 
\end{verbatim}
\end{codesize}

\noindent
Even Obsidian programs 
that never explicitly uses a Push array will also be represented by this 
datatype.

Note that the {\tt Par} constructor, 
the {\em parallel for loop}, could potentially introduce nesting, which would lead 
to {\em nested data-parallelism}. We do not compile nested data parallelism 
into CUDA, and right now this is guaranteed by taking care not to introduce 
any nesting in the library functions provided. Some of the 
simpler cases of nestedness should be possible to take care of quite easily. 
For example, one extra level of nesting could be done by sequential execution 
in each thread of the GPU; using sequential computations per thread has 
been shown to be beneficial~\cite{OLAMARCUS}. But for the general case 
of arbitrary nesting, some method of flattening is needed. 
We also assume that both {\tt Allocate} and {\tt Synchronize} occur only at the top level in objects of type {\tt Program}.

Now, a Push array is a function in continuation passing style
coupled with a length.

\begin{codesize}
\begin{verbatim}
data ArrayP a = ArrayP (P (UWordE, a)) Word32
\end{verbatim}
\end{codesize}

There is a function that takes an array and turns it into a Push array, 
called {\tt push}. This function is defined for both Pull and Push arrays:

\begin{codesize}
\begin{verbatim} 
class Pushable a where 
  push :: a e -> ArrayP e

instance Pushable ArrayP where 
  push = id

instance Pushable Array where   
  push (Array ixf n) =
     ArrayP (\func -> Par (\i -> func (i,(ixf i))) n) n 
\end{verbatim}
\end{codesize}

Going in the other direction, from a Push array to a Pull array, is a 
costly operation; it involves writing all the elements to GPU memory 
followed by creating a Pull array that represents reading them. 
The task of writing intermediate values to memory 
has traditionally been up to the {\tt sync} operation in Obsidian.
Therefore, in this version, {\tt sync} is overloaded to operate on both Pull and 
Push arrays.  This means that the 
{\tt sync} operation can be used both on arrays of type {\tt Array} 
and of type {\tt ArrayP}. The result type, however, is always {\tt Array}.

When a Push array is synced, it is applied to a continuation that writes the
elements into a named array in memory. The name to use is obtained through the 
{\tt Kernel} monad.

\begin{codesize}
\begin{verbatim}
targetArray :: Scalar a => Name -> (UWordE,Exp a) -> Program
targetArray n (i,a) = Assign n i a 
\end{verbatim}
\end{codesize}

After applying the Push array to {\tt targetArray <name>}, the {\tt sync}
operation proceeds by storing away a representation of the program that 
computes the array called {\tt <name>}; it returns a Pull array 
that reads elements from that same array.

Now we have seen enough of the implementation of Push arrays to be able 
to look at some operations. Earlier, we saw that the
array concatenation function
{\tt conc} on Pull arrays leads to inefficient code. The Push 
version of this operation, called {\tt concP} can be implemented 
as follows:

\begin{codesize}
\begin{verbatim}
concP :: (Pushable arr1,
          Pushable arr2) => (arr1 a, arr2 a) -> ArrayP a     
concP (arr1,arr2) = 
  ArrayP (\func -> f func
                   *>* 
                   g (\(i,a) -> func (fromIntegral n1 + i,a)))
         (n1+n2)
  where 
     ArrayP f n1 = push arr1
     ArrayP g n2 = push arr2
\end{verbatim}
\end{codesize}

\noindent
The function {\tt concP} takes two arrays, that can be Push or Pull arrays, 
and concatenates them into a single Push array. It does so by creating a 
sequential program, using the {\tt *>*} operator for {\tt Program} sequential
composition. An example use of this combinator has already been displayed in 
the {\tt catArrayPs} example.

%If the {\tt catArrays} kernel is reimplemented using {\tt concP}, CUDA code 
%without conditionals is generated. The new version of {\tt catArrays}, 
%called {\tt catArrayPs} is:

%\begin{codesize}
%\begin{verbatim}
%catArrayPs :: (Array IntE, Array IntE) -> Kernel (ArrayP Int)
%catArrayPs = pure concP
%\end{verbatim}
%\end{codesize}

%And the CUDA code it generates is:

%\begin{codesize}
%\begin{verbatim}
%__global__ void catArrayPs(int *input0,int *input1,int *result0){
%  unsigned int tid = threadIdx.x;
%  unsigned int bid = blockIdx.x;
%  
%  result0[((bid*32)+tid)] = input0[((bid*16)+tid)];
%  result0[((bid*32)+(16+tid))] = input1[((bid*16)+tid)];
%  
%}
%\end{verbatim}
%\end{codesize}
%__global__ void catArrayPs(int *input0,
%                           int *input1,
%                           int *result0){
%  unsigned int tid = threadIdx.x;
%  unsigned int bid = blockIdx.x;
%  
%  result0[((bid*64)+tid)] = input0[((bid*32)+tid)];
%  result0[((bid*64)+(32+tid))] = input1[((bid*32)+tid)];
%}

The {\tt zippUnpair} example shows a drawback similar to that of {\tt catArrays} using Pull arrays. 
In this case, the problem is that the {\tt unpair} function introduces 
a conditional that takes different paths depending on odd or even {\em thread id}. 
A Push array implementation of the {\tt unpair} operation called {\tt unpairP}
can be given as follows:

\begin{codesize}
\begin{verbatim}
unpairP :: Pushable arr => arr (a,a) -> ArrayP a 
unpairP arr =  ArrayP (\k -> f (everyOther k)) (2 * n)
  where 
    ArrayP f n = push arr

everyOther :: ((UWordE, a) -> Program ()) 
              -> (UWordE, (a,a)) -> Program ()
everyOther f  = \(ix,(a,b)) -> f (ix * 2,a) *>* f (ix * 2 + 1,b)  
\end{verbatim}
\end{codesize}

\noindent
Just like {\tt concP}, this function takes either a Push or Pull array 
as input, and produces a Push array as result.

Rewriting the example from earlier using {\tt unpairP} gives:

\begin{codesize}
\begin{verbatim}
zippUnpairP :: (Array IntE, Array IntE) -> Kernel (ArrayP IntE) 
zippUnpairP = pure (unpairP . zipp)
\end{verbatim}
\end{codesize}

\noindent
In this case, the generated code looks as follows:

\begin{codesize}
\begin{verbatim}
__global__ void zippUnpairP(int *input0,int *input1,int *result0){
  unsigned int tid = threadIdx.x;
  unsigned int bid = blockIdx.x;

  result0[((bid*64)+(tid*2))] = input0[((bid*32)+tid)];
  result0[((bid*64)+((tid*2)+1))] = input1[((bid*32)+tid)];
}
\end{verbatim}
\end{codesize}

\noindent
Again, we get CUDA code that uses half as many threads as the inefficient 
version, but all threads are occupied at all times. This uses
the resources more efficiently.

%\subsection{Push arrays and Obsidian}

Being able to generate the kind of code that we have just seen is something we have desired for a long time. We believe that 
Push arrays are an important tool for obtaining high performance kernels. The results 
in section~\ref{sec:MARY} bear this out.

%Section~\ref{sec:BENEFITS} showed some examples where Push arrays 
%was helpful. Lets look at how the operations on Push arrays used there 
%are implemented. First there was {\tt concP}:

%The function {\tt concP} can be used to concatenate any arrays that we can 
%perform the {\tt push} function on. For example both input arrays could be 
%normal Pull arrays. Concatenating the two arrays uses sequential composition 
%of programs, {\tt *>*}, which is implemented using the {\tt ProgramSeq} 
%constructor.

%The next function used in the examples was {\tt unpairP}

%\subsection {Benefits of Push Arrays}
%\label{sec:BENEFITS}

%in section~\ref{sec:Drawbacks} some drawbacks of Pull arrays where 
%highlighted. If we use Push arrays we can improve the generated code 
%in those examples. Push arrays here have type {\tt ArrayP} and operations 
%that operate on Push arrays or give Push arrays as result end their names
%with a capital p.

%In the case of array concatenation using Push arrays we get:

%\begin{codesize}
%\begin{verbatim}
%catArrayPs :: (Array (Exp Int), Array (Exp Int)) 
%           -> Kernel (ArrayP (Exp Int))
%catArrayPs = pure concP
%\end{verbatim}
%\end{codesize}
%(arr1,arr2) = 
%   return$ concP (arr1, arr2)

%The function {\tt push} is used to turn a Pull array into a Push array. 
%Now, precisely the code we desired can be generated:

%\pagebreak

%\begin{codesize}
%\begin{verbatim}
%__global__ void catArrayPs(int *input0,
%                           int *input1,
%                           int *result0){
%  unsigned int tid = threadIdx.x;
%  unsigned int bid = blockIdx.x;

%  result0[((bid*64)+tid)] = input0[((bid*32)+tid)];
%  result0[((bid*64)+(32+tid))] = input1[((bid*32)+tid)];

%}
%\end{verbatim}
%\end{codesize}

%The next example uses {\tt unpair}:

%\subsection { Thoughts about Push arrays in Obsidian }

%\input{applications}
\section{Application}
\label{sec:MARY}

\subsection{Sorting on a GPU}
In this section, we introduce combinators that express patterns
of computation on whole arrays, and show their application to the
development of fast sorting kernels.
By kernel, we mean a computation that is performed by multiple
threads, each performing the same computation, in a single block.
The computation is performed entirely on the GPU, operating on
a short array, which has been placed into shared memory.
On the GPU on which we perform measurements, the maximum number
of threads in a single block is 512. Thus, we will build and benchmark
a sequence of
kernels that sort 512 inputs.
Our first kernels are implemented using one thread per array element.
Next, we show how Push arrays allow us to move to having each
thread operate on two array elements, giving a substantial performance
improvement.

Sorting kernels are typically used as building blocks in larger
programs to sort much larger sequences of inputs. In section~\ref{sec:benchmarks},
we show how to build a sorter for large arrays from small building blocks, including
small kernels for sorting and merging that are generated from Obsidian.
Our small kernels are constructed in the form of sorting and merging {\em networks},
building on Batcher's bitonic merger~\cite{Batcher} and
on the periodic balanced merger~\cite{PeriodicBalanced}.
We chose also to implement the large sorter used  in benchmarking the
small kernels as a sorting network. However, large
sorters that are not themselves sorting networks (with typical examples being radix sort and quicksort) often call small sorting networks when they need to sort small arrays during their execution.
Thus, small, fast sorting kernels have a variety of uses.

\subsection{Describing Batcher's bitonic merger}

The bitonic merger is typically presented as a recursive
construction and we have earlier explored ways to describe
and analyse it in both (our) Ruby and in Lava~\cite{sortsRuby,LavaSorter}.
Here, we consider iterative descriptions using similar combinators.

Figure~\ref{fig:bitonicMerger} illustrates the merger for 16 inputs.
Data flows from left to right.
The vertical lines indicate components that operate on two array elements, placing the minimum onto the lower (abstract) wire, and the maximum onto the other output
of the component.
The leftmost {\em stage} operates (for $n=16$) on elements that are 
$8$ apart. the next stage deals with elements that are $4$ apart, and so on.

\begin{figure}
\centering
\includegraphics[scale=0.25]{bitonic}
\caption{A diagram of a 16 input bitonic merging network, using
a style that is standard in the literature. Note
that in each stage containing 8 min/max or comparator components, all
8 operate on independent parts of the input and so can proceed in parallel.
}
\label{fig:bitonicMerger}
\end{figure}

We introduce a combinator {\tt ilv1}, for {\em interleave}, that captures this pattern.
{\tt ilv1 i f g} applies {\tt f} to elements {\small $2^i$} apart,
producing the output at the lower of the two input indices; it applies {\tt g}
to the same pairs of elements, producing the output on the upper index.
Defining {\tt stage i} to be {\tt ilv1 i min max},
the four stages in the diagram are simply
{\tt stage} applied to{\small  $3$, $2$ $1$} and {\small $0$}.
The definition of {\tt ilv1} makes use of the fact that
flipping the bit $i$ of an index (using
the function {\tt flipBit}) gives the index of the element
that will be combined with it using the functions{\tt f} and {\tt g}.
The decision about whether to apply {\tt f} or {\tt g} is made
by looking at the value of bit $i$.
As we shall see, the use of the Obsidian {\tt ifThenElse}
produces conditionals in the resulting CUDA.
\begin{codesize}
\begin{verbatim}
lowBit :: Int -> UWordE -> Exp Bool
lowBit i ix = (ix .&. bit i) ==* 0

flipBit :: Bits a => Int -> a -> a
flipBit = flip complementBit

ilv1 :: Choice a => 
        Int -> (b -> b-> a) -> (b -> b -> a) -> 
        Array b -> Array a
ilv1 i f g arr = Array ixf (len arr)
  where
    ixf ix = let l = arr ! ix
                 r = arr ! newix
                 newix = flipBit i ix
             in (ifThenElse (lowBit i ix) (f l r) (g l r))
\end{verbatim}
\end{codesize}
\noindent
Expressing {\tt ilv1} using bit-flipping may seem strange, but it has
the advantage that it actually applies the desired pattern of
computation repeatedly over larger input arrays.
Now, for {\small $2^n$} inputs, a Haskell list containing the $n$ calls
of this interleave combinator are built:
\begin{codesize}
\begin{verbatim}
bmerge :: Int -> [Array IntE -> Array IntE]
bmerge n = [istage (n-i) | i <- [1..n]]
  where istage i = ilv1 i min max
\end{verbatim}
\end{codesize}
\noindent
Finally, the {\tt compose} function makes each element of the list
into a kernel (using {\tt map pure}) and places a {\tt sync} between
each kernel (using {\tt composeS}).
\begin{codesize}
\begin{verbatim}
compose :: (Scalar a) => 
           [Array (Exp a) -> Array (Exp a)] 
           -> Array (Exp a) -> Kernel (Array (Exp a))
compose = composeS . map pure

runm k = putStrLn$ CUDA.genKernel "bitonicMerge" 
         (compose (bmerge k)) (namedArray "inp" (2^k))
\end{verbatim}
\end{codesize}
\noindent
Note that {\tt bmerge k} works on inputs of length {\small }$2^{k+j}$, for $j > 0$, applying the merger to sub-sequences of length {\small $2^k$}.
The CUDA code for {\tt bmerge 4} on $16$ inputs
(with some newlines inserted) is
\begin{codesize}
\begin{verbatim}
*Main> runm 4
__global__ void bitonicMerge(int *input0,int *result0){
  unsigned int tid = threadIdx.x;
  unsigned int bid = blockIdx.x;
  extern __shared__ unsigned char sbase[];
  (( int *)sbase)[tid] 
     = ((tid&8)==0) 
      ? min(input0[((bid*16)+tid)],input0[((bid*16)+(tid^8))]) 
      : max(input0[((bid*16)+tid)],input0[((bid*16)+(tid^8))]);
  __syncthreads();
  (( int *)(sbase + 64))[tid] 
     = ((tid&4)==0) 
      ? min((( int *)sbase)[tid],(( int *)sbase)[(tid^4)]) 
      : max((( int *)sbase)[tid],(( int *)sbase)[(tid^4)]);
  __syncthreads();
  (( int *)sbase)[tid] 
     = ((tid&2)==0) 
      ? min((( int *)(sbase+64))[tid],(( int *)(sbase+64))[(tid^2)]) 
      : max((( int *)(sbase+64))[tid],(( int *)(sbase+64))[(tid^2)]);
  __syncthreads();
  (( int *)(sbase + 64))[tid] 
     = ((tid&1)==0) 
      ? min((( int *)sbase)[tid],(( int *)sbase)[(tid^1)]) 
      : max((( int *)sbase)[tid],(( int *)sbase)[(tid^1)]);
  __syncthreads();
  result0[((bid*16)+tid)] = (( int *)(sbase+64))[tid];
\end{verbatim}
\end{codesize}

%% ** TODO Note that bmerge k not only works on inputs of length $2^{k+j}$,
%% for $j > 0$, applying the merger to sub-sequences of length $2^k$.

\subsection{Modifying the bitonic merger}

The bitonic merger for which we have just generated a kernel is known
to sort so-called bitonic sequences, which include sequences whose first
half is sorted in one direction and whose second half is sorted in the other
direction. This fact can be used to build the well-known bitonic sorting
network. However, a GPU implementation typically needs to check, for each comparator,
whether or not it should sort upwards or downwards, see for instance
the simple CUDA implementation shown in Appendix A.
We choose here to modify the merger so that it sorts two concatenated
sequences that are sorted in the {\em same} direction.
We do this by using a well-known trick, reversing half of the input to the merger. It turns out that
we can also reverse the same half of the output of the first stage of
the network, without affecting overall behaviour.
The resulting network, {\tt tmerge}, shown in Figure~\ref{fig:mixedMerger},
encourages us to develop a new combinator to describe the characteristic
V-shaped pattern that results in the first stage.
The combinator is modelled on {\tt ilv1}. The only difference
is that the ``partner'' of an index is found not by flipping bit $i$, but
by flipping bits $0$ to $i$, using function {\tt flipLSBsTo}. The implementation
of {\tt vee1} is got from that for {\tt ilv1} by replacing the call of
{\tt flipBit} by one of {\tt flipLSBsTo} (and we could also have chosen
to make a more generic function that is parameterised on this {\em partner} function).
\begin{codesize}
\begin{verbatim}
flipLSBsTo :: Int -> UWordE -> UWordE
flipLSBsTo i = (`xor` (oneBits (i+1)))

vee1 :: Choice a => 
        Int -> (b -> b-> a) -> (b -> b -> a) -> 
        Array b -> Array a
vee1 i f g arr = Array ixf (len arr)
  where
    ixf ix = let l = arr ! ix
                 r = arr ! newix
                 newix = flipLSBsTo i ix
             in (ifThenElse (lowBit i ix) (f l r) (g l r))
\end{verbatim}
\end{codesize}

\begin{figure}
\centering
\includegraphics[scale=0.25]{mixed}
\caption{16 input merging network, {\em tmerge}. The first stage
is made with a new combinator that we call vee, while
the remaining stages are as in the bitonic merger.
This network sorts an input that consists of two half-sized
sorted sequences, giving the opportunity to build a tree-shaped
sorting network.}
\label{fig:mixedMerger}
\end{figure}

\begin{codesize}
\begin{verbatim}
tmerge :: Int -> [Array IntE -> Array IntE]
tmerge n = vstage (n-1): [istage (n-i) | i <- [2..n]]
  where
    vstage i = vee1 i min max 
    istage i = ilv1 i min max
\end{verbatim}
\end{codesize}

Now that we have a merger that sorts sub-sequences containing two concatenated sorted sequences, it
is easy to make a tree of them.
The list of kernels to be composed now becomes
\begin{codesize}
\begin{verbatim}
tsort1 :: Int -> [Array IntE -> Array IntE]
tsort1 n = concat [tmerge i | i <- [1..n]]
\end{verbatim}
\end{codesize}
\noindent
The resulting sorting network is shown in Figure~\ref{fig:mixedsorter}.

\begin{figure}
\centering
\includegraphics[scale=0.33]{mmixedsorter}
\caption{A 16-input sorter made from a tree of {\em tmerge} mergers. 8 2-input mergers
feed 4 4-input mergers, followed  by 2 8-input and one 16-input merger.
A sorter on 8 inputs is shaded, as is a merger on 8 inputs, above it. 
}
\label{fig:mixedsorter}
\end{figure}
\noindent
The following call writes the resulting CUDA to file {\tt tsort1.cu}
and this is the first {\em generated} CUDA kernel whose performance is measured in section~\ref{sec:benchmarks}.

\begin{codesize}
\begin{verbatim}
runs1 k
  = writeFile "tsort1.cu" $ CUDA.genKernel "tsort1" 
      (compose (tsort1 k)) (namedArray "inp" (2^k))
\end{verbatim}
\end{codesize}

The generated code
% for $8$ inputs is shown in Appendix A. It 
uses
one thread per array element (just as the {\tt bitonicMerge}
kernel shown above did). The next step is to move to using Push as
well as Pull arays, so as to be able to generate more efficient code
from essentially the same sorter construction.

\subsection{New combinator implementations using Push arrays}

%% ** TODO make code for ilv2 more comprehensible :)

It is in combining the results of the {\tt f} and {\tt g} functions
that we run into difficulty using just Pull arrays.
Earlier, we saw how to use Push arrays to implement {\tt concP},
which concatenates two arrays.
Here, we use exactly the same approach to make a new version of
the {\em interleave} combinator. The results of applying
the {\tt f}s and {\tt gs} are combined into a Push array, in the right
order.
\pagebreak
%% ** TODO More explanation needed.
\begin{codesize}
\begin{verbatim}
ixMap :: (UWordE -> UWordE) -> ArrayP a -> ArrayP a 
ixMap f (ArrayP p n) = ArrayP (ixMap' f p) n

ixMap' :: (UWordE -> UWordE) 
         -> P (UWordE, a)
         -> P (UWordE, a) 
ixMap' f p = \g -> p (\(i,a) -> g (f i,a))

insertZero :: Int -> UWordE -> UWordE
insertZero 0 a = a `shiftL` 1
insertZero i a 
  = a + (a .&. fromIntegral (complement (oneBits i :: Word32)))

ilv2 :: Choice b => 
        Int -> (a -> a -> b) -> (a -> a -> b) -> 
        Array a -> ArrayP b
ilv2 i f g (Array ixf n) 
   = ArrayP (\k -> app a5 k *>* app a6 k) n
  where
    n2 = n `div` 2
    a1 = Array (ixf . left) (n-n2)
    a2 = Array (ixf . right) n2
    a3 = zipWith f a1 a2
    a4 = zipWith g a1 a2
    a5 = ixMap left (push a3)
    a6 = ixMap right (push a4)
    left = insertZero i
    right = flipBit i  . left
    app (ArrayP f _) a = f a
\end{verbatim}
\end{codesize}
\noindent
This new combinator can now replace {\tt ilv1} in the bitonic
merger, giving a kernel that runs considerably faster. We will use
that kernel to build a large sorter later.

The implemenation of {\tt vee2} is almost identical to
that of {\tt ilv2}, with {\tt flipBit i}
replaced by {\tt flipLSBsTo} as before (so that, again, one would in
fact make a more generic function for building such combinators).
Now, we just need to replace the {\tt ilv1} and {\tt vee1} combinators
in the tree sorter with {\tt ilv2} and {\tt vee2} , to get a verrsion that uses half as many threads:
\begin{codesize}
\begin{verbatim}
tmerge2 :: Int -> [Array IntE -> ArrayP IntE]
tmerge2 n = vstage (n-1) : [ istage (n-i) | i <- [2..n]]
  where
    vstage i = vee2 i min max
    istage i = ilv2 i min max

tsort2 :: Int -> [Array IntE -> ArrayP IntE]
tsort2 n = concat [tmerge2 i | i <- [1..n]]
\end{verbatim}
\end{codesize}
As we shall see in section~\ref{sec:benchmarks}, the resulting code
is significantly faster. This is because it uses one thread per two array elements,
and the code no longer contains any conditionals.

In order to go faster still, we resort to building a different sorting network, of
exactly the same size as the bitonic sorter, but based instead on the balanced period merger of Dowd et al~\cite{PeriodicBalanced}.
This involves the introduction of one new combinator that can be seen as a mixture
of the {\tt ilv} and {\tt vee} combinators already introduced.

\subsection{A sorter built from the balanced periodic merger}
First, we note that the balanced periodic merger contains multiple
uses of the now familiar vee-shaped pattern, see Figure~\ref{fig:periodicMerger}.
\begin{figure}
\centering
\includegraphics[scale=0.25]{balanced}
\caption{16 input periodic balanced merging network}
\label{fig:periodicMerger}
\end{figure}

\begin{codesize}
\begin{verbatim}
bpmerge2 :: Int -> [Array IntE -> ArrayP IntE]
bpmerge2 n = [vstage (n-i) | i <- [1..n]]
  where vstage i = vee2 i min max
\end{verbatim}
\end{codesize}

Now Dowd et al proved that the balanced periodic merger sorts two {\em interleaved} sorted sequences. So, taking an iterative view of the resulting sorter,
we want to build a tree of mergers as before, but the smaller mergers should
be interleaved, rather than operating on adjacent sub-sequences.
There should be one merger on the right hand end of the network; left of
that, there should be two mergers that operate on the odd and even elements,
and each of them should in turn be fed by two interleaved mergers, and so.
The sorter is illustrated, for $16$ inputs in Figure~\ref{fig:vsorter}.
Just to the left of the final balanced merger, one of the two interleaved
mergers is shown using dotted lines. It operates on a completely different set of inputs from the other 8-input merger in the same part of the tree.

\begin{figure*}
\centering
\includegraphics[scale=0.5]{vsorter}
\caption{A sorter based on the idea that the periodic
balanced merger network sorts two interleaved sorted sequences.
It consists of two half-sized sorters, one working on the
odd elements of the input and one on the even, followed by
the balanced merger. The diagram indicates using dotted lines the
balanced merger that is the final (rightmost) part of
one of the half-size sorters.}
\label{fig:vsorter}
\end{figure*}

The most straightforward way to give an iterative description of
this sorter is
to introduce a new combinator that is a combination of
{\tt ilv} and {\tt vee}.
The only thing that we need to change is the {\em partner} function.
This time we will flip not the least significant bits from position $0$ to
position $i$ but from position $i$ to $i+j$.

\begin{codesize}
\begin{verbatim}
-- flip bits from position i to position i+j inclusive
flipBitsFrom :: Bits a => Int -> Int -> a -> a
flipBitsFrom i j a = a `xor` (fromIntegral mask)
  where
    mask = (oneBits (j + 1):: Word32) `shiftL` i

ilvVee1 :: Choice a => 
           Int -> Int -> 
           (b -> b-> a) -> (b -> b -> a) -> 
           Array b -> Array a
ilvVee1 i j f g arr = Array ixf (len arr)
  where
    ixf ix = let l = arr ! ix
                 r = arr ! newix
                 newix = flipBitsFrom i j ix
             in (ifThenElse (lowBit (i+j) ix) (f l r) (g l r))
\end{verbatim}
\end{codesize}

\begin{codesize}
\begin{verbatim}
ilvVee2 :: Choice b => Int -> Int -> 
           (a -> a -> b) -> (a -> a -> b) -> 
           Array a -> ArrayP b
ilvVee2 i j f g (Array ixf n) 
  = ArrayP (\k -> app a5 k *>* app a6 k) n
  where
    n2 = n `div` 2
    a1 = Array (ixf . left) (n-n2)
    a2 = Array (ixf . right) n2
    a3 = zipWith f a1 a2
    a4 = zipWith g a1 a2
    a5 = ixMap left (push a3)
    a6 = ixMap right (push a4)
    left = insertZero (i+j)
    right = flipBitsFrom i j . left
    app (ArrayP f _) a = f a
\end{verbatim}
\end{codesize}

For both variants of the combinator, we simply add to the {\tt ilv} definitions
a new {\tt Int} parameter, {\tt j}, and replace {\tt flipBit i} by {\tt flipBitsFrom i j}. We also insert the zero bit (when calculating the left index)
at position {\tt i + j} rather than just at position {\tt i}.
{\tt ilvVee} is a generalisation of both {\tt ilv} and {\tt vee}.
{\tt ilvVee i 0} has the same behaviour as {\tt ilv i}, and {\tt ilvVee 0 (j-1)}
is the same as {\tt vee j}. The {\tt i} parameter controls the degree
of interleaving, and the {\tt j} parameter controls the size of
the vee-shaped blocks.

For $16$ inputs, the parameters
to {\tt ilvVee} that describe the periodic merger on the right of the construction
are $i=0$ (for no interleaving) paired with $3$, $2$, $1$ and $0$
for the decreasing size of the vee-shaped blocks.
Next, to the left, the mergers are interleaved ($i=1$) and there are three
stages with
vee-shaped blocks of decreasing size ($j = 2, 1, 0$), see Figure~\ref{fig:vsorter}.
The following code gives an iterative description of the construction
for {\small $2^n$} inputs:

\begin{codesize}
\begin{verbatim}
vsort :: Int -> Array IntE -> Kernel (Array IntE)
vsort n = composeS . map pure $ [istage (n-i) (i-j) 
                                | i <- [1..n], j <- [1..i]]
  where istage i j = ilvVee2 i j min max
\end{verbatim}
\end{codesize}

\noindent
The resulting generated code uses one thread per
$2$ indices.

\subsection{Measuring performance of the generated kernels}\label{sec:benchmarks}
\begin{figure}
\begin{codesize}
\begin{verbatim}
  unsigned int arrayLength = 1 << LOG_L_SIZE;
  unsigned int diff = LOG_L_SIZE - LOG_S_SIZE;
  unsigned int blocks = arrayLength / S_SIZE;
  unsigned int threads = S_SIZE / 2;

  sortSmall<<<blocks, threads,4096>>>(din,din);

  for(int i = 0 ; i < diff ; i += 1){ 
    vSwap<<<blocks/2,threads*2,0>>>(din,din,(1<<i)*S_SIZE);

    for(int j = i-1; j >= 0; j -= 1)
      iSwap<<<blocks/2,threads*2,0>>>(din,din,(1<<j)*S_SIZE);

    bmergeSmall<<<blocks,threads,4096>>>(din,din);}
\end{verbatim}
\end{codesize}
\caption{CUDA code for our large sorter. {\tt sortSmall} and {\tt bmergeSmall}
are replaced by Obsidian-generated kernels in the experiments. {\tt vSwap} and {\tt iSwap} are handwritten CUDA kernels
that perform one column of compare and swap operations in the vee and ilv shapes respectively, and are
parameterised on the stride. (Our generated kernels have fixed input and output size.)}
\label{fig:CUDAsort}
\end{figure}
We have measured the performance of generated 512-input sorting and merging kernels by plugging them into a larger sorter written in CUDA.
The sorter has exactly the structure shown in Figure~\ref{fig:mixedsorter}.
Figure~\ref{fig:newmixedsorter} shows the location of smaller sorters and mergers, and of vSwap and iSwap kernels for $16$ inputs.
Larger sorters simply have more columns of mergers, each preceded by a vSwap and a number of iSwaps.
The overall structure of the resulting CUDA code is shown in Figure~\ref{fig:CUDAsort}.
Because {\tt iSwap} is used repeatedly, we also wrote kernels
corresponding to compositions of two and three of them (avoiding
memory accesses between the columns).
That is, we replaced the loop containing {\tt iSwap} above by
\pagebreak
\begin{codesize}
\begin{verbatim}
for(int j = i-1; j >= 0; j -= 3){ 
  if (j==0) 
    iSwap<<<blocks/2,threads*2,0>>>(din,din,(1<<j)*S_SIZE);
    else 
    {if (j==1) 
       iSwap2<<<blocks/4,threads*2,0>>>(din,din,(1<<j)*S_SIZE);
       else 
       iSwap3<<<blocks/8,threads*2,0>>>(din,din,(1<<j)*S_SIZE);}}
\end{verbatim}
\end{codesize}
\noindent
The Obsidian implementation and all code for examples in the paper are available at \url{http://www.cse.chalmers.se/~joels/expressive.html}.

\begin{figure}
\centering
\includegraphics[scale=0.33]{newmixedSorter}
\caption{
This diagram shows the tree-shaped sorting network that we saw earlier, but with 4 input sorting and merging kernels indicated. This is the structure of the network that we have used to implement large sorters, in which the small kernels having {\small $2^9 = 512$} inputs in all cases. For {\small $2^{24}$} inputs overall, the resulting sorter
has one column of small sorters and $24-9=15$ of small mergers.}
%% both ofs in the above sentence should remain :)
\label{fig:newmixedsorter}
\end{figure}

Table~\ref{fig:table} shows the performance figures for the large sorter
for $5$ different $512-$input small sorter kernels.
{\tt tsort1} is defined above, and is a variant of bitonic sort
that does not require {\tt if} statements to determine the direction
of sorting of pairs, as all comparison operations place the minimum at the
same index as the lower input.
{\tt tsort2} is the same construction, but built with the combinators {\tt ilv2}
and {\tt vee2} that result in the use of one thread per two indices.
{\tt vsort}, the fastest kernel, uses a generalisation of those
two combinators, and again uses one thread per two indices.
{\tt vsort1} is the same construction as {\tt vsort}, but with {\tt ilvVee2} replaced
by {\tt ilvVee1}, and so using one thread per index.
As a reference small kernel implementation, we include a simple hand-coded bitonic sort
that uses one thread per index (see code in Appendix).
None of the kernels has been subject to optimisations related to warp size.

\input{table1.tex}

We also recorded the GPU time spent in the calls to the small sorters alone (using the NVIDIA CUDA Visual profiler), see Table~\ref{tab:table2}.
The {\tt tsort1} and {\tt vsort1} kernels, which are built using only pull arrays, 
(and which have one index per thread in the generated code) are
noticeably slower than those that have two indices per thread.
The generation of the latter kernels is made possible by the introduction of 
push arrays.

\input{table2.tex}

Although none of our generated kernels is optimised (for example with respect to warp size), 
their performance is, nevertheless, very good. We are working on automated warp size-related 
optimisations. It would also be interesting to explore the fusion of adjacent columns of 
comparators in the small kernels; omitting a {\tt sync} would cause this fusion to happen
but we also need to modify the threading behaviour (doubling the number of indices per thread).

%The final version of the paper will contain benchmarking against a highly optimised sorter kernel obtained from our colleagues in the computer graphics group at Chalmers. For now, we emphasise that the {\tt vsort} kernel is so fast that we expect it to compare favourably. Given the simplicity of its description in Obsidian, we feel that this is a good way to write fast kernels.

%\input{discussion}

\section {Discussion}

%% ** TODO. Mention CSE. warp-related opt.

Push arrays form a new approach to array representation in DSELs.
We do not know of similar approaches in the literature, despite
the fact that the notions of demand and data flow
may feel familiar to the reader who considers Pull and Push arrays.
The addition of Push arrays to Obsidian seems highly beneficial. With 
this new feature, the user gains finer grained control over the code generated
and the resulting CUDA kernels perform considerably better than before.
This 
was illustrated in the series of sorters explored in section~\ref{sec:MARY}.
The performance of {\tt vsort} is sufficiently good that it
can be used as a first phase in a larger sorter (written in CUDA) that
can sort 16M elements in 96 ms, while an i7-920 CPU takes
around 2740 ms. Further speed improvements look possible, both in the coordination code
and in the kernels. An obvious next step would be to investigate the generation of the {\tt iSwap} and
{\tt vSwap} kernels from Obsidian. (This is not currently possible because of assumptions that we made about the interfaces to kernels and about how {\em thread ids} are used. We will look into ways to relax
our assumptions.)

The series of kernels also illustrates how the use of combinators brings a form of reuse, and makes design exploration easier. Our experience of using similar combinators in the Lava hardware
description language~\cite{LavaSorter} was that a relatively small set of combinators went a long way. So, although we introduced three
combinators here, {\tt ilv}, {\tt vee} and {\tt ilvVee}, which
includes the other two, we do not believe that every new kernel development exercise would demand a completely new set of combinators.
We expect to provide the user with a well-documented set of combinators,
so that users can get access to this style of programming without having
to develop their own combinators, and without having to think too much
about bit-hacking.
The bit-manipulation approach chosen to define our combinators automatically
created functions that apply to sub-sequences of the input that
are of an appropriate length.

In this paper, we made combinators for the special case of two input, two
output operations (built from two two-input funtions that we typically called
{\tt f} and {\tt g}). 
This approach should be generalised to deal with blocks that have $2^k$ inputs
and outputs.
Also, we made a compound combinator from {\tt ilv} and {\tt vee},
but generalising to more than two input components would allow
for composing combinators, and indeed for recursive descriptions
that could be unrolled. Then, ignoring {\tt syncs}, a recursive description of
{\tt vsort} could be something like 
\begin{codesize}
\begin{verbatim}
vsortR 0 = id
vsortR n = bpmergeR n . ilv2 1 (vsortR (n-1))
\end{verbatim}
\end{codesize}
\noindent
It would then be necessary to optimise the code generated from
multiple applications of {\tt ilv2 1}, for example, whereas here
we have forced the user to figure out both the unrolling and the combinations.
Moving to more general combinators would also give the opportunity to
provide predefined combinators that capture more of the commonly used threading patterns (for instance $k$ indices per thread rather than the $1$ and $2$ shown here).

The integration of Push arrays into Obsidian raises some new questions.
Previously, there was a direct correspondence between the length 
of an array and the number of threads used to compute it, which
allowed the user to write an initial program without worrying about
threads at all, and then to tweak the Obsidian program if he was not satisfied
with the threading behaviour of the resulting kernel. 
Now, as can be seen in the {\tt catArrayPs} example and in the sorters, this correspondence
can be broken using Push arrays. The {\tt catArrayPs} example and two
of the sorters use half as many threads as the 
number of elements. For users who are very concerned about the speed of
the generated kernels, getting this control through using Push arrays in
a particular pattern is clearly a good thing. But adding a second, different
way to control thread use in the generated code certainly complicates matters, and further case studies are needed to confirm that the complication pays off.

The addition of Push arrays also adds the possibility to include potentially 
unsafe operations in Obsidian, for example by writing multiple array elements to the same index, or by discarding elements.
This new expressiveness will have to be carefully controlled. On the positive side, it offers the possibility to encode functions like {\tt filter} from Haskell
that
are simply not expressible using only Pull arrays. Being able to implement {\tt filter} would make programming kernels in obsidian feel much more like programming in Haskell -- a welcome loosening of the strait-jacket.
Once that is done, it will be time to develop a very simple coordination language to allow programming of entire GPU applications that make use of
the kind of small kernel building blocks developed here.

\appendix
\section{Appendix}
\begin{codesize}
\begin{verbatim}
__device__ inline void swap(int & a, int & b)
{
   int tmp = a;
   a = b;
   b = tmp;
}

__global__ static void bitonicSort(int * values, int *results)
{
   extern __shared__ int shared[];

   const unsigned int tid = threadIdx.x;
   const unsigned int bid = blockIdx.x;

   // Copy input to shared mem.
   shared[tid] = values[(bid*NUM) + tid];

   __syncthreads();

   // Parallel bitonic sort.
   for (unsigned int k = 2; k <= NUM; k *= 2)
   { 
     // bitonic merge
     for (unsigned int j = k / 2; j>0; j /= 2)
       {
           unsigned int ixj = tid ^ j;

           if (ixj > tid)
           {
               if ((tid & k) == 0)
               {
                   if (shared[tid] > shared[ixj])
                   {
                       swap(shared[tid], shared[ixj]);
                   }
               }
               else
               {
                   if (shared[tid] < shared[ixj])
                   {
                       swap(shared[tid], shared[ixj]);
                   }
               }
           }
           __syncthreads();
       }
   }

   // Write result.
   results[(bid*NUM) + tid] = shared[tid];
}
\end{verbatim}
\end{codesize}

\acks
This research has been funded by the Swedish Foundation for Strategic Research (which funds
the RAW FP Project) and
by the Swedish Research Council.

% We recommend abbrvnat bibliography style.

%\bibliographystyle{abbrvnat}
%\bibliography{paper}

% The bibliography should be embedded for final submission.

\begin{thebibliography}{}
\softraggedright

\bibitem[Axelsson et~al.(2011)Axelsson, Claessen, Sheeran, Svenningsson,
  Engdal, and Persson]{FELDSPAR}
E.~Axelsson, K.~Claessen, M.~Sheeran, J.~Svenningsson, D.~Engdal, and
  A.~Persson.
\newblock {T}he {D}esign and {I}mplementation of {F}eldspar an {E}mbedded
  {L}anguage for {D}igital {S}ignal {P}rocessing.
\newblock In \emph{Proceedings of the 22nd international conference on
  Implementation and application of functional languages}, IFL'10, pages
  121--136, Berlin, Heidelberg, 2011. Springer-Verlag.

\bibitem[Batcher(1968)]{Batcher}
K.~E. Batcher.
\newblock Sorting {N}etworks and {T}heir {A}pplications.
\newblock In \emph{Proc. AFIPS Spring Joint Computer Conference 32}, pages
  307--314, 1968.

\bibitem[Billeter et~al.(2009)Billeter, Olsson, and Assarsson]{OLAMARCUS}
M.~Billeter, O.~Olsson, and U.~Assarsson.
\newblock Efficient stream compaction on wide {SIMD} many-core architectures.
\newblock In \emph{Proc. Conf. on High Performance Graphics}, HPG '09, pages
  159--166. ACM, 2009.

\bibitem[Chakravarty et~al.(2011)Chakravarty, Keller, Lee, McDonell, and
  Grover]{ACCELERATE}
M.~M. Chakravarty, G.~Keller, S.~Lee, T.~L. McDonell, and V.~Grover.
\newblock Accelerating {H}askell array codes with multicore {GPU}s.
\newblock In \emph{Proc. sixth workshop on {D}eclarative {A}spects of
  {M}ulticore {P}rogramming}, DAMP '11, pages 3--14. ACM, 2011.

\bibitem[Claessen(1999)]{POORKOEN}
K.~Claessen.
\newblock A poor man's concurrency monad.
\newblock \emph{J. Funct. Program.}, 9\penalty0 (3):\penalty0 313--323, 1999.

\bibitem[Claessen et~al.(2001)Claessen, Sheeran, and Singh]{LavaSorter}
K.~Claessen, M.~Sheeran, and S.~Singh.
\newblock The {D}esign and {V}erification of a {S}orter {C}ore.
\newblock In \emph{Proc. Int. Conf. on Correct Hardware Design and Verification
  Methods (CHARME)}, volume 2144 of \emph{Springer LNCS}, pages 355--369, 2001.

\bibitem[Dowd et~al.(1989)Dowd, Perl, and Larry~Rudolph]{PeriodicBalanced}
M.~Dowd, Y.~Perl, and M.~S. Larry~Rudolph.
\newblock The {P}eriodic {B}alanced {S}orting {N}etwork.
\newblock \emph{Journal of the ACM}, 36:\penalty0 738--757, 1989.

\bibitem[Elliott(2003)]{PAN}
C.~Elliott.
\newblock Functional {I}mages.
\newblock In \emph{The Fun of Programming}, ``Cornerstones of Computing''
  series. Palgrave, Mar. 2003.

\bibitem[Elliott et~al.(2003)Elliott, Finne, and de~Moor]{COMPILEEDSL}
C.~Elliott, S.~Finne, and O.~de~Moor.
\newblock Compiling embedded languages.
\newblock \emph{Journal of Functional Programming}, 13\penalty0 (2), 2003.

\bibitem[{J}oel {S}vensson(2011)]{JSLIC}
{J}oel {S}vensson.
\newblock {O}bsidian: {GPU} {K}ernel {P}rogramming in {H}askell.
\newblock Technical Report 77L, {C}omputer {S}cience and {E}nginering,
  {C}halmers {U}niversity of {T}echnology, {G}othenburg, 2011.
\newblock Thesis for the degree of Licentiate of Philosophy.

\bibitem[Keller et~al.(2010)Keller, Chakravarty, Leshchinskiy, Peyton~Jones,
  and Lippmeier]{REPA}
G.~Keller, M.~M. Chakravarty, R.~Leshchinskiy, S.~Peyton~Jones, and
  B.~Lippmeier.
\newblock Regular, shape-polymorphic, parallel arrays in {H}askell.
\newblock In \emph{Proceedings of the 15th ACM SIGPLAN international conference
  on Functional programming}, ICFP '10, pages 261--272, New York, NY, USA,
  2010. ACM.

\bibitem[Larsen(2011)]{BARRACUDA}
B.~Larsen.
\newblock Simple optimizations for an applicative array language for graphics
  processors.
\newblock In \emph{Proc. sixth workshop on {D}eclarative {A}spects of
  {M}ulticore {P}rogramming}, DAMP '11, pages 25--34. ACM, 2011.

\bibitem[Mainland and Morrisett(2010)]{NIKOLA}
G.~Mainland and G.~Morrisett.
\newblock Nikola: embedding compiled {GPU} functions in {H}askell.
\newblock In \emph{Proc. third ACM Symposium on Haskell}. ACM, 2010.

\bibitem[Sheeran(1989)]{sortsRuby}
M.~Sheeran.
\newblock Describing {B}utterfly {N}etworks in {R}uby.
\newblock In \emph{Functional Programming}, pages 182--205. Springer Workshops
  in Computing, 1989.

%\bibitem[Smith et~al.(2009)Smith, Jones]{smith02}
%P. Q. Smith, and X. Y. Jones. ...reference text...

\end{thebibliography}

\end{document}