### 2013-09-30 src2/expressive/push.tex

\section{Push Arrays}

In order to improve low level control for the programmer, {\em Push
arrays} are added to Obsidian. The old Pull arrays are still available,
along with the new array type.

Some operations, typically involving taking arrays apart,
are easily described using Pull arrays, giving
efficient code. In those cases, using a Push array would add complexity in the 
implementation for no performance benefit.
Other operations cannot be implemented efficiently
with Pull arrays, but Push arrays then provide the solution.
This duality is apparent when looking at operations on Pull arrays such as
{\tt halve} and {\tt conc} (for concatenate). 
The {\tt halve} function is efficient since it introduces no diverging 
conditionals. The {\tt conc} function, on the other hand, introduces conditionals.
Concatening two arrays using the {\tt concP} combinator, implemented on Push arrays, 
allows us to generate the desired code:

\begin{codesize}
\begin{verbatim}
catArrayPs :: (Array IntE, Array IntE) -> Kernel (ArrayP Int)
catArrayPs = pure concP
\end{verbatim}
\end{codesize}

\begin{codesize}
\begin{verbatim}
__global__ void catArrayPs(int *input0,int *input1,int *result0){
  unsigned int tid = threadIdx.x;
  unsigned int bid = blockIdx.x;

  result0[((bid*32)+tid)] = input0[((bid*16)+tid)];
  result0[((bid*32)+(16+tid))] = input1[((bid*16)+tid)];

}
\end{verbatim}
\end{codesize}

\noindent
Compared to the CUDA code for {\tt catArrays}, this kernel 
uses only $16$ threads instead of $32$. At each step of the computation, all 
the threads are fully busy doing exactly the same thing, which is the preferred 
mode of execution on the target platform.

%The performance effects of these unwanted conditionals, compared to code generated 
%using the new Push arrays, will be illustrated in section~\ref{sec:MARY}.

%The {\tt halve} operation has appeared previously and the {\tt pair} 
%operation is its inverse. Implementing the {\tt pair} operation using 
%Pull arrays is very easy and also efficient:

%\begin{codesize}
%\begin{verbatim}
%pair :: Array a -> Array (a,a)
%pair (Array ixf n) = Array (\ix -> (ixf (ix*2),ixf (ix*2+1))) n'
%  where 
%    n' = n `div` 2 
%\end{verbatim}
%\end{codesize}

%It is efficient because it introduces no diverging conditionals. There is
%just a tiny bit of extra arithmetic in the indexing function. On the 
%other hand {\tt unpair}, as we saw, introduced an unnecessary and 
%inefficient conditional. It turns out that {\tt unpair} can be 
%efficiently implemented on a Push array.

\subsection {What are Push Arrays?}

The idea behind Push arrays is to have a way to describe where 
elements are supposed to end up. In some sense, a Push array produces 
a collection of Index/Value pairs. %This makes a Push array a lot more 
%powerful than a Pull array. 
This makes Push arrays complementary to Pull arrays. For example, it 
is possible for a Push array to output several elements at the same index 
(which we probably need to control carefully). Push arrays should permit 
us to provide more expressive operations on arrays to the user, including 
an operation similar to Haskell's {\tt filter} on lists. Here, we consider 
a different advantage of adding push arrays: finer control over patterns of 
thread use in generated code.

%% Will it really give nondeterminism or will the last one written win, or some such?

A Push array consists of three parts: a function in continuation passing style, 
a {\tt Program} datatype and an array datatype.

\begin{codesize}
\begin{verbatim}
type P a = (a -> Program) -> Program 
\end{verbatim}
\end{codesize}

\noindent
For another example of using continuations and a more complete
description of their meaning and application, see~\cite{POORKOEN}.

The {\tt Program} datatype has now been adopted as Obsidian's internal 
representation of CUDA programs.

\begin{codesize}
\begin{verbatim}
data Program
  = Skip
  | forall a. Scalar a => Assign Name UWordE (Exp a) 
  | Par (UWordE -> Program) Word32   
  | Allocate Name Word32 Type 
  | Synchronize 
  | ProgramSeq Program 
               Program 
\end{verbatim}
\end{codesize}

\noindent
Even Obsidian programs 
that never explicitly uses a Push array will also be represented by this 
datatype.

Note that the {\tt Par} constructor, 
the {\em parallel for loop}, could potentially introduce nesting, which would lead 
to {\em nested data-parallelism}. We do not compile nested data parallelism 
into CUDA, and right now this is guaranteed by taking care not to introduce 
any nesting in the library functions provided. Some of the 
simpler cases of nestedness should be possible to take care of quite easily. 
For example, one extra level of nesting could be done by sequential execution 
in each thread of the GPU; using sequential computations per thread has 
been shown to be beneficial~\cite{OLAMARCUS}. But for the general case 
of arbitrary nesting, some method of flattening is needed. 
We also assume that both {\tt Allocate} and {\tt Synchronize} occur only at the top level in objects of type {\tt Program}.

Now, a Push array is a function in continuation passing style
coupled with a length.

\begin{codesize}
\begin{verbatim}
data ArrayP a = ArrayP (P (UWordE, a)) Word32
\end{verbatim}
\end{codesize}

There is a function that takes an array and turns it into a Push array, 
called {\tt push}. This function is defined for both Pull and Push arrays:

\begin{codesize}
\begin{verbatim} 
class Pushable a where 
  push :: a e -> ArrayP e

instance Pushable ArrayP where 
  push = id

instance Pushable Array where   
  push (Array ixf n) =
     ArrayP (\func -> Par (\i -> func (i,(ixf i))) n) n 
\end{verbatim}
\end{codesize}

Going in the other direction, from a Push array to a Pull array, is a 
costly operation; it involves writing all the elements to GPU memory 
followed by creating a Pull array that represents reading them. 
The task of writing intermediate values to memory 
has traditionally been up to the {\tt sync} operation in Obsidian.
Therefore, in this version, {\tt sync} is overloaded to operate on both Pull and 
Push arrays.  This means that the 
{\tt sync} operation can be used both on arrays of type {\tt Array} 
and of type {\tt ArrayP}. The result type, however, is always {\tt Array}.

When a Push array is synced, it is applied to a continuation that writes the
elements into a named array in memory. The name to use is obtained through the 
{\tt Kernel} monad.

\begin{codesize}
\begin{verbatim}
targetArray :: Scalar a => Name -> (UWordE,Exp a) -> Program
targetArray n (i,a) = Assign n i a 
\end{verbatim}
\end{codesize}

After applying the Push array to {\tt targetArray <name>}, the {\tt sync}
operation proceeds by storing away a representation of the program that 
computes the array called {\tt <name>}; it returns a Pull array 
that reads elements from that same array.

Now we have seen enough of the implementation of Push arrays to be able 
to look at some operations. Earlier, we saw that the
array concatenation function
{\tt conc} on Pull arrays leads to inefficient code. The Push 
version of this operation, called {\tt concP} can be implemented 
as follows:

\begin{codesize}
\begin{verbatim}
concP :: (Pushable arr1,
          Pushable arr2) => (arr1 a, arr2 a) -> ArrayP a     
concP (arr1,arr2) = 
  ArrayP (\func -> f func
                   *>* 
                   g (\(i,a) -> func (fromIntegral n1 + i,a)))
         (n1+n2)
  where 
     ArrayP f n1 = push arr1
     ArrayP g n2 = push arr2
\end{verbatim}
\end{codesize}

\noindent
The function {\tt concP} takes two arrays, that can be Push or Pull arrays, 
and concatenates them into a single Push array. It does so by creating a 
sequential program, using the {\tt *>*} operator for {\tt Program} sequential
composition. An example use of this combinator has already been displayed in 
the {\tt catArrayPs} example.

%If the {\tt catArrays} kernel is reimplemented using {\tt concP}, CUDA code 
%without conditionals is generated. The new version of {\tt catArrays}, 
%called {\tt catArrayPs} is:

%\begin{codesize}
%\begin{verbatim}
%catArrayPs :: (Array IntE, Array IntE) -> Kernel (ArrayP Int)
%catArrayPs = pure concP
%\end{verbatim}
%\end{codesize}

%And the CUDA code it generates is:

%\begin{codesize}
%\begin{verbatim}
%__global__ void catArrayPs(int *input0,int *input1,int *result0){
%  unsigned int tid = threadIdx.x;
%  unsigned int bid = blockIdx.x;
%  
%  result0[((bid*32)+tid)] = input0[((bid*16)+tid)];
%  result0[((bid*32)+(16+tid))] = input1[((bid*16)+tid)];
%  
%}
%\end{verbatim}
%\end{codesize}
%__global__ void catArrayPs(int *input0,
%                           int *input1,
%                           int *result0){
%  unsigned int tid = threadIdx.x;
%  unsigned int bid = blockIdx.x;
%  
%  result0[((bid*64)+tid)] = input0[((bid*32)+tid)];
%  result0[((bid*64)+(32+tid))] = input1[((bid*32)+tid)];
%}

The {\tt zippUnpair} example shows a drawback similar to that of {\tt catArrays} using Pull arrays. 
In this case, the problem is that the {\tt unpair} function introduces 
a conditional that takes different paths depending on odd or even {\em thread id}. 
A Push array implementation of the {\tt unpair} operation called {\tt unpairP}
can be given as follows:

\begin{codesize}
\begin{verbatim}
unpairP :: Pushable arr => arr (a,a) -> ArrayP a 
unpairP arr =  ArrayP (\k -> f (everyOther k)) (2 * n)
  where 
    ArrayP f n = push arr

everyOther :: ((UWordE, a) -> Program ()) 
              -> (UWordE, (a,a)) -> Program ()
everyOther f  = \(ix,(a,b)) -> f (ix * 2,a) *>* f (ix * 2 + 1,b)  
\end{verbatim}
\end{codesize}

\noindent
Just like {\tt concP}, this function takes either a Push or Pull array 
as input, and produces a Push array as result.

Rewriting the example from earlier using {\tt unpairP} gives:

\begin{codesize}
\begin{verbatim}
zippUnpairP :: (Array IntE, Array IntE) -> Kernel (ArrayP IntE) 
zippUnpairP = pure (unpairP . zipp)
\end{verbatim}
\end{codesize}

\noindent
In this case, the generated code looks as follows:

\begin{codesize}
\begin{verbatim}
__global__ void zippUnpairP(int *input0,int *input1,int *result0){
  unsigned int tid = threadIdx.x;
  unsigned int bid = blockIdx.x;

  result0[((bid*64)+(tid*2))] = input0[((bid*32)+tid)];
  result0[((bid*64)+((tid*2)+1))] = input1[((bid*32)+tid)];
}
\end{verbatim}
\end{codesize}

\noindent
Again, we get CUDA code that uses half as many threads as the inefficient 
version, but all threads are occupied at all times. This uses
the resources more efficiently.

%\subsection{Push arrays and Obsidian}

Being able to generate the kind of code that we have just seen is something we have desired for a long time. We believe that 
Push arrays are an important tool for obtaining high performance kernels. The results 
in section~\ref{sec:MARY} bear this out.

%Section~\ref{sec:BENEFITS} showed some examples where Push arrays 
%was helpful. Lets look at how the operations on Push arrays used there 
%are implemented. First there was {\tt concP}:

%The function {\tt concP} can be used to concatenate any arrays that we can 
%perform the {\tt push} function on. For example both input arrays could be 
%normal Pull arrays. Concatenating the two arrays uses sequential composition 
%of programs, {\tt *>*}, which is implemented using the {\tt ProgramSeq} 
%constructor.

%The next function used in the examples was {\tt unpairP}

%\subsection {Benefits of Push Arrays}
%\label{sec:BENEFITS}

%in section~\ref{sec:Drawbacks} some drawbacks of Pull arrays where 
%highlighted. If we use Push arrays we can improve the generated code 
%in those examples. Push arrays here have type {\tt ArrayP} and operations 
%that operate on Push arrays or give Push arrays as result end their names
%with a capital p.

%In the case of array concatenation using Push arrays we get:

%\begin{codesize}
%\begin{verbatim}
%catArrayPs :: (Array (Exp Int), Array (Exp Int)) 
%           -> Kernel (ArrayP (Exp Int))
%catArrayPs = pure concP
%\end{verbatim}
%\end{codesize}
%(arr1,arr2) = 
%   return$ concP (arr1, arr2)

%The function {\tt push} is used to turn a Pull array into a Push array. 
%Now, precisely the code we desired can be generated:

%\pagebreak

%\begin{codesize}
%\begin{verbatim}
%__global__ void catArrayPs(int *input0,
%                           int *input1,
%                           int *result0){
%  unsigned int tid = threadIdx.x;
%  unsigned int bid = blockIdx.x;

%  result0[((bid*64)+tid)] = input0[((bid*32)+tid)];
%  result0[((bid*64)+(32+tid))] = input1[((bid*32)+tid)];

%}
%\end{verbatim}
%\end{codesize}

%The next example uses {\tt unpair}:

%\subsection { Thoughts about Push arrays in Obsidian }