### 2013-09-30 src2/expressive/table.tex

% table

\begin{figure}
\begin{center}
  \begin{tabular}{  l | r | r}
    %\hline
    Kernel               & Threads   & ms / grid   \\ \hline
    bitonic\_sort(CUDA)  & 512       & 30.28       \\      
    tsort1               & 512       & 15.82       \\
    tsort2               & 256       & 11.64       \\ 
    vsort1               & 512       & 14.96       \\ 
    vsort                & 256       &  9.33       \\ 
    %\hline
  \end{tabular}
\end{center}
\label{fig:table}
\caption{Table is showing the time it takes for a grid of 32678 blocks of 
threads to execute.Each grid was launched 1000 times, the total execution time 
was then divided by 1000 to get the listed figures. The GPU used is a NVIDIA GTX480} 
\end{figure}