### 2013-09-30 src2/expressive/table2.tex

% table

\begin{table}
\begin{center}
  \begin{tabular}{| l | r | r |   }

    \hline
                              & time & percent \\ \hline
    bitonic(CUDA)             & 30203 & 32.07\\
    tsort1                    & 15823 & 19.72  \\
    vsort1                       & 14955 & 18.76 \\ 
    tsort2                      & 11562 & 15.37    \\ 
    vsort                       & 9228 &  12.67 \\ 
    bitonic (NVIDIA SDK)        & 23815 & 6.55 \\
    \hline
  \end{tabular}
\end{center}
\label{tab:table2}
\caption{time: GPU time (in $\mu$s) spent in calls to small sorter kernels 
in the initial phase of the sorter for $2^{24}$ inputs. percent: percentage 
of total GPU time spent in the small sort kernel.
The final line shows the numbers for the bitonic code (for large sorts) that 
is distributed with the the NVIDIA SDK. Its structure also starts with a phase 
of small sorting kernels. It takes 274 ms to sort $2^{24}$ key-value pairs on 
an NVIDIA GTX480 GPU. It could be sped up by using our {\tt vsort} kernel, 
and also by (hand) fusing single ``column'' kernels as we did with {\tt iSwap}, 
although the need to calculate directions of comparators in the bitonic network 
would complicate this. Our CUDA code is simpler because all comparators
point in the same direction in the sorter construction.
} 
\end{table}

%% bitonic 30203
%% tsort1 15823
%% tsort2 11562
%% vsort1 14955
%% vsort 9228