### 2013-09-30 src2/csort/paper.tex

%-----------------------------------------------------------------------------
%
%               Template for sigplanconf LaTeX Class
%
% Name:         sigplanconf-template.tex
%
% Purpose:      A template for sigplanconf.cls, which is a LaTeX 2e class
%               file for SIGPLAN conference proceedings.
%
% Guide:        Refer to "Author's Guide to the ACM SIGPLAN Class,"
%               sigplanconf-guide.pdf
%
% Author:       Paul C. Anagnostopoulos
%               Windfall Software
%               978 371-2316
%               paul@windfall.com
%
% Created:      15 February 2005
%
%-----------------------------------------------------------------------------

\documentclass[]{sigplanconf}

% The following \documentclass options may be useful:
%
% 10pt          To set in 10-point type instead of 9-point.
% 11pt          To set in 11-point type instead of 9-point.
% authoryear    To obtain author/year citation style instead of numeric.

\usepackage{amsmath}

\usepackage{graphicx}
\usepackage{placeins} 
\usepackage{url}
\usepackage[T1]{fontenc}
\usepackage{tikz}
\usepackage{fancyvrb}

\begin{document}

\exclusivelicense
\conferenceinfo{FHPC~'13}{September 23 2013, Boston, MA, USA} 
\copyrightyear{2013} 
\copyrightdata{978-1-4503-2381-9/13/09} 
\doi{2502323.2502325}

\titlebanner{DRAFT}        % These are ignored unless
\preprintfooter{Counting Sort}   % 'preprint' option specified.

%\title{Efficient Counting Sort Implementations using an Embedded GPU Programming Language}
%\title{Implementing Counting Sort for GPUs using an Embedded Language}
\title{Counting and Occurrence Sort for GPUs using an Embedded Language}
%\subtitle{Subtitle Text, if any}

\authorinfo{Josef Svenningsson \and Bo Joel Svensson \and Mary Sheeran}
           {Dept, of Computer Science and Engineering \\ 
            Chalmers University of Technology}
           {\{josefs, joels, ms\}@chalmers.se}

\maketitle

\begin{abstract}
This paper investigates two sorting algorithms: counting sort and a
variation, occurrence sort,  which also removes duplicate elements, 
and examines their suitability for running on the GPU. The duplicate 
removing variation
turns out to have a natural functional, data-parallel implementation
which makes it particularly interesting for GPUs.

The algorithms are implemented in Obsidian, a high-level domain
specific language for GPU programming.

Measurements show that our implementations in many cases outperform
the sorting algorithm provided by the library Thrust. Furthermore, occurrence
sort is another factor of two faster than
ordinary counting sort. We conclude that counting sort is an important
contender when considering sorting algorithms for the GPU, and that 
occurrence sort is highly preferable when applicable. We also show 
that Obsidian can produce very competitive code.
\end{abstract}

%% \category{CR-number}{subcategory}{third-level}
%% <<<<<<< HEAD

%% %% \terms
%% %% term1, term2

%% %% \keywords
%% %% keyword1, keyword2
%% \category{D.3.2}{Programming Languages}{Language Classifications}[Applicative (functional) languages; Concurrent, distributed, and parallel languages]
%% \category{D.3.4}{Programming Languages}{Processors}[Code generation]

%% %\terms
%% %term1, term2
%% \terms
%% Languages, Performance
%% =======
\category{D.3.2}{Programming Languages}{Language Classifications}[Applicative (functional) languages; Concurrent, distributed, and parallel languages]
\category{D.3.4}{Programming Languages}{Processors}[Code generation]
%% \terms
%% term1, term2
%% >>>>>>> 3b812338c4f7649982b17fb45df51f13d4285484

%\keywords
%keyword1, keyword2
\keywords
%% <<<<<<< HEAD
%% Data parallelism, array programming, embedded language
%% =======
Sorting, embedded language

%% >>>>>>> 3b812338c4f7649982b17fb45df51f13d4285484

\input{intro.tex}
\input{Algorithm.tex}
\input{GPUs.tex}
\input{Obsidian.tex} 
\input{ObsCudaAlgo.tex}
\input{benchmarks.tex}
\input{discussion.tex}

%\appendix
%\section{Appendix Title}

%This is the text of the appendix, if you need one.

\acks
This research has been funded by the Swedish Foundation for
Strategic Research (which funds the Resource Aware Functional 
Programming (RAW FP) Project) and by the Swedish Research Council.

\bibliographystyle{abbrvnat}
%% \bibliography{paper}

\begin{thebibliography}{18}
\providecommand{\natexlab}[1]{#1}
\providecommand{\url}[1]{\texttt{#1}}
\expandafter\ifx\csname urlstyle\endcsname\relax
  \providecommand{\doi}[1]{doi: #1}\else
  \providecommand{\doi}{doi: \begingroup \urlstyle{rm}\Url}\fi

\bibitem[wik()]{wikipedia}
Wikipedia article: Counting sort.
\newblock \url{http://en.wikipedia.org/wiki/Counting_sort}.

\bibitem[Chakravarty et~al.(2011)Chakravarty, Keller, Lee, McDonell, and
  Grover]{Acc}
M.~M. Chakravarty, G.~Keller, S.~Lee, T.~L. McDonell, and V.~Grover.
\newblock {Accelerating Haskell array codes with multicore GPUs}.
\newblock In \emph{Proc. of the sixth workshop on Declarative aspects of
  multicore programming}, DAMP '11. ACM, 2011.

\bibitem[Claessen et~al.(2012)Claessen, Sheeran, and Svensson]{PUSH}
K.~Claessen, M.~Sheeran, and B.~J. Svensson.
\newblock {E}xpressive {A}rray {C}onstructs in an {E}mbedded {GPU} {K}ernel
  {P}rogramming {L}anguage.
\newblock In \emph{Proceedings of the 7th workshop on Declarative aspects and
  applications of multicore programming}, DAMP '12, 2012.

\bibitem[Elliott et~al.(2003)Elliott, Finne, and de~Moor]{ELLIJFP}
C.~Elliott, S.~Finne, and O.~de~Moor.
\newblock Compiling embedded languages.
\newblock \emph{Journal of Functional Programming}, 13\penalty0 (2), 2003.
\newblock URL \url{http://conal.net/papers/jfp-saig/}.

\bibitem[Harris et~al.(2007)Harris, Sengupta, and Owens]{LargeScan}
M.~Harris, S.~Sengupta, and J.~D. Owens.
\newblock Parallel {P}refix {S}um ({S}can) with {CUDA}.
\newblock In H.~Nguyen, editor, \emph{GPU Gems 3}. Addison Wesley, 2007.

\bibitem[{J}oel {S}vensson(2011)]{JSLIC}
{J}oel {S}vensson.
\newblock {O}bsidian: {GPU} {K}ernel {P}rogramming in {H}askell.
\newblock Technical Report 77L, {C}omputer {S}cience and {E}nginering,
  {C}halmers {U}niversity of {T}echnology, {G}othenburg, 2011.
\newblock Thesis for the degree of Licentiate of Philosophy.

\bibitem[Karras(2012)]{Karras}
T.~Karras.
\newblock {Maximizing Parallelism in the Construction of BVHs, Octrees, and k-d
  Trees}.
\newblock In \emph{Proc. of the Fourth ACM SIGGRAPH / Eurographics conference
  on High-Performance Graphics}, EGGH-HPG'12. Eurographics Association, 2012.

\bibitem[Knuth(1998)]{Knuth}
D.~E. Knuth.
\newblock \emph{The art of computer programming, volume 3: (2nd ed.) sorting
  and searching}.
\newblock Addison Wesley Longman Publishing Co., Inc., Redwood City, CA, USA,
  1998.
\newblock ISBN 0-201-89685-0.

\bibitem[Kolonias et~al.(2011)Kolonias, Voyiatzis, Goulas, and Housos]{CSORT}
V.~Kolonias, A.~G. Voyiatzis, G.~Goulas, and E.~Housos.
\newblock {Design and implementation of an efficient integer count sort in CUDA
  GPUs}.
\newblock \emph{{Concurrency And Computing: Practice And Experiance}},
  23\penalty0 (18), Dec. 2011.

\bibitem[Krueger et~al.(2011)Krueger, Grund, Jaeckel, Zeier, and
  Plattner]{REMOVEDUPS}
J.~Krueger, M.~Grund, I.~Jaeckel, A.~Zeier, and H.~Plattner.
\newblock {Applicability of GPU Computing for Efficient Merge in In-Memory
  Databases}.
\newblock In \emph{The Second International Workshop on Accelerating Data
  Management Systems using Modern Processor and Storage Architectures (ADMS
  11)}, 2011.
\newblock \url{http://www.adms-conf.org/p19-KRUEGER.pdf}.

\bibitem[Kulkarni and Newton(2013)]{FPCDSL}
A.~Kulkarni and R.~R. Newton.
\newblock {E}mbrace, {D}efend, {E}xtend: {A} {M}ethodology for {E}mbedding
  {P}reexisting {DSL}s, 2013.
\newblock Functional Programming Concepts in Domain-Specific Languages
  (FPCDSL'13).

\bibitem[Mainland and Morrisett(2010)]{NIKOLA}
G.~Mainland and G.~Morrisett.
\newblock {Nikola: embedding compiled {GPU} functions in Haskell}.
\newblock \emph{SIGPLAN Not.}, 45\penalty0 (11), 2010.

\bibitem[{NVIDIA}({\natexlab{a}})]{CUDA}
{NVIDIA}.
\newblock {CUDA C Programming Guide}, {\natexlab{a}}.
\newblock URL
  \url{http://docs.nvidia.com/cuda/cuda-c-programming-guide/index.html}.

\bibitem[{NVIDIA}({\natexlab{b}})]{KEPLER}
{NVIDIA}.
\newblock {NVIDIAs Next Generation CUDA Compute Architecture: Kepler GK110},
  {\natexlab{b}}.
\newblock URL
  \url{http://www.nvidia.com/content/PDF/kepler/NVIDIA-Kepler-GK110-Architecture-Whitepaper.pdf}.

\bibitem[{NVIDIA}({\natexlab{c}})]{THRUST}
{NVIDIA}.
\newblock {NVIDIA Thrust Library}, {\natexlab{c}}.
\newblock URL \url{https://developer.nvidia.com/thrust}.

\bibitem[Olsson et~al.(2012)Olsson, Billeter, and Assarsson]{Olsson}
O.~Olsson, M.~Billeter, and U.~Assarsson.
\newblock Clustered deferred and forward shading.
\newblock In \emph{Proc. of the Fourth ACM SIGGRAPH / Eurographics conference
  on High-Performance Graphics}, EGGH-HPG'12. Eurographics Association, 2012.

\bibitem[Sintorn and Assarsson(2008)]{sintorn}
E.~Sintorn and U.~Assarsson.
\newblock Fast parallel {GPU}-sorting using a hybrid algorithm.
\newblock \emph{Journal of Parallel and Distributed Computing}, 68\penalty0
  (10), 2008.

\bibitem[Sklansky(1960)]{Sklansky}
J.~Sklansky.
\newblock {Conditional Sum Addition Logic}.
\newblock \emph{Trans. IRE}, EC-9\penalty0 (2), June 1960.

\end{thebibliography}

%% % The bibliography should be embedded for final submission.

%% \begin{thebibliography}{}
%% \softraggedright

%% \bibitem[Smith et~al.(2009)Smith, Jones]{smith02}
%% P. Q. Smith, and X. Y. Jones. ...reference text...

%% \end{thebibliography}

\end{document}