### 2013-09-30 src2/embarbb/conclusion.tex

\section{Conclusion} 
We have shown that Intel's Array Building Blocks (ArBB) provides
an interface that is well suited to functional programming. 
The programs are quite close to mathematical specifications, in the style of NESL~\cite{NESL}. We have only just completed the embedding of the part of ArBB that deals with nested vectors, and we need to tackle many more case studies.
In our case studies using dense vectors, ArBB seems to do a good job of efficiently using parallel
hardware resources -- both cores and vector units.
By
embedding ArBB, we can, with little implementation effort, provide
quite an attractive data parallel programming language in Haskell.
Our benchmarks are preliminary and small, but they show very good performance.
Once we have completed the ArBB embedding, we will have an interesting platform on which to
experiment with and develop new programming idioms that exploit the fact
that we have a data parallel programming language embedded in an
expressive, strongly typed host language.
We feel that work in this area (as distinct from implementation methods)
is overdue. We would be happy to receive suggestions
for interesting case studies or collaborations.