### 2013-09-30 src2/embarbb/future.tex

\section{Future Work} 
\label{sec:FutureWork}

%The Haskell - EmbArBB interface needs to be improved. 
%Currently there is a 
%transfer of data into ArBB before a function is executed and the results 
%are directly transferred back into Haskell.
%This is not ideal when several ArBB functions are composed, as this
%will lead to unnecessary copying of data into and out of ArBB.
%One possible solution would be to have the {\tt ArBB} monad keep track of what arrays have been bound 
%to ArBB variables and stored into ArBB memory.
%Another problem with the interface is the way {\tt map} works. Currently 
%{\tt map} takes an identifier pointing out a captured function. These
%captured functions are maintained by the {\tt ArBB} monad, which means that 
%either a function that maps some other function over data needs to be defined 
%in the {\tt ArBB} monad, or it needs to be passed the function as an argument 
%at capture time. Neither of those alternatives is desirable and new solutions
%will have to be found.

%One thought is to 
%also let the {\tt ArBB} monad keep track of what arrays have been bound 
%to ArBB variables and stored into ArBB memory. Currently there is a 
%transfer of data into ArBB before a function is executed and the results 
%are directly transferred back into Haskell. This is limiting in case 
%the programmer wants to execute several ArBB functions in sequence, each 
%on the result of the previous function. In this scenario data will be 
%copied back and forwards needlessly between function executions in the current 
%implementation.

The C++ embedding of ArBB allows for dense containers of structs in some cases. 
The operations on vectors supplied by the ArBB virtual machine are exclusively 
over vectors of scalar types. So the C++ embedding must be performing a AOS to SOA 
(Array of Struct to Struct of Array) transformation. The Haskell embedding 
does not implement any similar transformation. This is an important addition 
that would for example make implementing functions on complex numbers easier.

We stress that this paper presents first steps in the implementation of 
EmbArBB. Our benchmarks, while promising, are very limited. We must devote 
effort to developing a suite of interesting, larger data parallel programs 
for use in benchmarking EmbArBB. It is particularly important to explore the 
nested vectors, and the effects of the limitation to one level of nesting 
in ArBB. Those parts of ArBB that support nested vectors seem to be less well 
developed than those supporting dense vectors, as evidenced by the sample 
applications distributed with ArBB, none of which uses nesting. We expect 
ArBB to become more complete, and perhaps we will be able to contribute 
interesting examples both in the C++ and Haskell embeddings.

Having exercised EmbArBB more thoroughly, we will assess the results of 
the benchmarking and experiments with programming idioms, and decide on future
research directions. We expect to focus on ways to provide users with an interface that is more 
functional in style than the current C++ oriented one. 
%ArBB has nested vectors that are used to implement nested data parallelism, that 
%is data parallelism with irregularly shaped data, in the style of
%NESL~\cite{NESL}. The Haskell embedding has so 
%far been focused entirely on the dense functionality of ArBB. 
%The next step will be to find ways to embed the nested operations too.
%Looking on the 
%nested operations and seeing how these can be embedded as well is next in line.

%GPUs are efficient on flat data parallel workloads, and approaches
%to GPU programming often offer only flat data parallelism.
%Although there is some promising recent work on nested data parallelism
%in a functional setting
%on GPUs~\cite{NestedGPU}, this area is generally less well explored.
%We are enticed by the idea of
%using EmbArBB as a 
%platform for exploration of GPU programming and nested data-parallelism.
%Implementing a GPU backend to EmbArBB would require applying 
%GPU specific optimisations to the EmbArBB AST (expression datatype). 
%The question of whether to perform the necessary optimisations
%at the EmbArBB level, or later, inside ArBB, is an interesting
%one that we would like to explore.

%However 
%if implementing a GPU backend on the EmbArBB level is the right thing to do 
%or if rather it should be implemented as part of the ArBB system? There 
%could be a point to explore in the setting of EmbArBB the space of different 
%optimisations and transformations needed to run the functions on a GPU before 
%doing it in the more complicated (perhaps) setting of full ArBB.