### 2013-09-30 src2/embarbb/paper.tex

\documentclass{sigplanconf}

% The following \documentclass options may be useful:
%
% 10pt          To set in 10-point type instead of 9-point.
% 11pt          To set in 11-point type instead of 9-point.
% authoryear    To obtain author/year citation style instead of numeric.

%\usepackage[none]{hyphenat}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{url}
\usepackage{graphicx}
\usepackage{placeins}
\usepackage{fancyvrb}

\begin{document}
\conferenceinfo{FHPC'12,} {September 15, 2012, Copenhagen, Denmark.}
\CopyrightYear{2012}
\copyrightdata{978-1-4503-1577-7/12/09}

\titlebanner{}        % These are ignored unless
\preprintfooter{}   % 'preprint' option specified.

% \title{}
\title{Parallel Programming in Haskell Almost for Free} 
\subtitle{an embedding of Intel's Array Building Blocks}

\authorinfo{Bo Joel Svensson}
           {Chalmers University of Technology}
           {joels@chalmers.se}
\authorinfo{Mary Sheeran}
           {Chalmers University of Technology}
           {ms@chalmers.se}

\maketitle

\begin{abstract}
\input{abstract.tex}
\end{abstract}

%\category{CR-number}{subcategory}{third-level}
\category{D.3.2}{Programming Languages}{Language Classifications}[Applicative (functional) languages; Concurrent, distributed, and parallel languages]
\category{D.3.4}{Programming Languages}{Processors}[Code generation]

\terms
Languages, Performance

\keywords
Data parallelism, array programming, dynamic compilation, embedded language

%\input{outline.tex}

\input{introduction.tex}
\input{related.tex} 
\input{programming.tex}
\input{implementation.tex}
\input{results.tex}

%\input{discussion}
\input{future.tex}
\input{conclusion.tex}

%\cite{joelryan}
%\cite{arbbpb}
%\cite{arbbcpp}
%\cite{arbbvm}
%\cite{ARBB2011}
%\cite{REPA}
%\appendix
%\section{Appendix Title}

%This is the text of the appendix, if you need one.

\acks

%** Note. Joel should thank Intel. We should ack funding from VR and SSF.
Svensson was first introduced to ArBB while on a three month internship at Intel
during 2011. Thanks go to Ryan R. Newton for inspiring supervision 
during that internship.
%while 
%working on Haskell bindings to ArBB at Intel as well as trying to understand 
%the intricate details of the inside of Data.Array.Accelerate.

This research has been funded by the Swedish Foundation for
Strategic Research (which funds the Resource Aware Functional 
Programming (RAW FP) Project) and by the
Swedish Research Council.

% We recommend abbrvnat bibliography style.

\bibliographystyle{abbrvnat}

% The bibliography should be embedded for final submission.

\begin{thebibliography}{}
\softraggedright
\bibitem[Axelsson et~al.(2010)Axelsson, Claessen, D{\'e}vai, Horv{\'a}th,
  Keijzer, Lyckeg{\aa}rd, Persson, Sheeran, Svenningsson, and
  Vajda]{FELDSPAR2010}
E.~Axelsson, K.~Claessen, G.~D{\'e}vai, Z.~Horv{\'a}th, K.~Keijzer,
  B.~Lyckeg{\aa}rd, A.~Persson, M.~Sheeran, J.~Svenningsson, and A.~Vajda.
\newblock Feldspar: {A} domain specific language for digital signal processing
  algorithms.
\newblock In \emph{8th ACM/IEEE International Conference on Formal Methods and
  Models for Codesign (MEMOCODE 2010)}, pages 169--178. IEEE Computer Society,
  2010.

\bibitem[Blelloch(1996)]{NESL}
G.~Blelloch.
\newblock Programming {P}arallel {A}lgorithms.
\newblock \emph{Communications of the ACM}, 39\penalty0 (3), 1996.

\bibitem[Chakravarty et~al.(2011)Chakravarty, Keller, Lee, McDonell, and
  Grover]{accelerate}
M.~M. Chakravarty, G.~Keller, S.~Lee, T.~L. McDonell, and V.~Grover.
\newblock {A}ccelerating {H}askell {A}rray {C}odes with {M}ulticore {GPU}s.
\newblock In \emph{Proceedings of the sixth workshop on Declarative aspects of
  multicore programming}, DAMP '11, pages 3--14, New York, NY, USA, 2011. ACM.
\newblock ISBN 978-1-4503-0486-3.
%\newblock \doi{10.1145/1926354.1926358}.
\newblock URL \url{http://doi.acm.org/10.1145/1926354.1926358}.

\bibitem[Elliott et~al.(2003)Elliott, Finne, and de~Moor]{ELLIJFP}
C.~Elliott, S.~Finne, and O.~de~Moor.
\newblock Compiling embedded languages.
\newblock \emph{Journal of Functional Programming}, 13\penalty0 (2), 2003.
\newblock URL \url{http://conal.net/papers/jfp-saig/}.

\bibitem[Gill(2009)]{Gill}
A.~Gill.
\newblock Type-{S}afe {O}bservable {S}haring in {H}askell.
\newblock In \emph{Proceedings of the 2009 ACM SIGPLAN Haskell Symposium},
  09/2009 2009.
\newblock URL
  \url{http://www.ittc.ku.edu/csdl/fpg/sites/default/files/Gill-09-TypeSafeRei%
fication.pdf}.

\bibitem[Hughes and Sheeran(2012)]{TFPIEPFP}
J.~Hughes and M.~Sheeran.
\newblock Teaching parallel functional programming at {C}halmers.
\newblock In \emph{presentation at Trends in Functional Programming in
  Education, Workshop associated with Conf. on Trends in Functional
  Programming, St. Andrews}, 2012.

\bibitem[Intel()]{arbbvm}
Intel.
\newblock {I}ntel(r) {A}rray {B}uilding {B}locks {V}irtual {M}achine
  {S}pecification.
\newblock \url{http://software.intel.com/sites/whatif/arbb/arbb_vm.pdf}.

\bibitem[Keller et~al.(2010)Keller, Chakravarty, Leshchinskiy, Peyton~Jones,
  and Lippmeier]{REPA}
G.~Keller, M.~M. Chakravarty, R.~Leshchinskiy, S.~Peyton~Jones, and
  B.~Lippmeier.
\newblock Regular, shape-polymorphic, parallel arrays in haskell.
\newblock In \emph{Proceedings of the 15th ACM SIGPLAN international conference
  on Functional programming}, ICFP '10, pages 261--272, New York, NY, USA,
  2010. ACM.
\newblock ISBN 978-1-60558-794-3.
%\newblock \doi{10.1145/1863543.1863582}.
\newblock URL \url{http://doi.acm.org/10.1145/1863543.1863582}.

\bibitem[Lippmeier et~al.(2012)Lippmeier, Chakravarty, Keller, Leshchinskiy,
  and Jones]{DPH}
B.~Lippmeier, M.~M.~T. Chakravarty, G.~Keller, R.~Leshchinskiy, and S.~P.
  Jones.
\newblock {W}ork {E}fficient {H}igher-{O}rder {V}ectorisation.
\newblock \url{http://www.cse.unsw.edu.au/~chak/papers/replicate-tr.pdf}, 2012.
\newblock ICFP'12.

\bibitem[Mainland and {M}orrisett(2010)]{NIKOLA}
G.~Mainland and G.~{M}orrisett.
\newblock Nikola: {E}mbedding {C}ompiled {GPU} {F}unctions in {H}askell.
\newblock In \emph{Proceedings of the third ACM Haskell symposium}, pages
  67--78. ACM, 2010.
\newblock ISBN 978-1-4503-0252-4.
\newblock URL \url{http://doi.acm.org/10.1145/1863523.1863533}.

\bibitem[Newburn et~al.(2011)Newburn, So, Liu, McCool, Ghuloum, Toit, Wang, Du,
  Chen, Wu, Guo, Liu, and Zhang]{ARBB2011}
C.~J. Newburn, B.~So, Z.~Liu, M.~McCool, A.~Ghuloum, S.~D. Toit, Z.~G. Wang,
  Z.~H. Du, Y.~Chen, G.~Wu, P.~Guo, Z.~Liu, and D.~Zhang.
\newblock Intel's array building blocks: A retargetable, dynamic compiler and
  embedded language.
\newblock In \emph{Proceedings of the 2011 9th Annual IEEE/ACM International
  Symposium on Code Generation and Optimization}, CGO '11, pages 224--235,
  Washington, DC, USA, 2011. IEEE Computer Society.
\newblock ISBN 978-1-61284-356-8.

\bibitem[Sheeran(2004)]{LavaMultipliers}
M.~Sheeran.
\newblock Generating fast multipliers using clever circuits.
\newblock In \emph{Int. Conf. on Formal Methods in Computer Aided Design
  (FMCAD)}, volume 3312 of \emph{LNCS}, pages 6--20, 2004.

\bibitem[Sheeran(2011)]{SheeranJFP}
M.~Sheeran.
\newblock Functional and dynamic programming in the design of parallel prefix
  networks.
\newblock \emph{J. Funct. Program.}, 21\penalty0 (1):\penalty0 59--114, 2011.

\bibitem[Svensson and Newton(2011)]{joelryan}
B.~J. Svensson and R.~Newton.
\newblock {P}rogramming {F}uture {P}arallel {A}rchitectures with {H}askell and
  {A}r{BB}.
\newblock \url{http://faspp.ac.upc.edu/faspp11/pdf/faspp11-final12.pdf}, 2011.
\newblock Presented at the workshop: {F}uture {A}rchitectural {S}upport for
  {P}arallel {P}rogramming (FASPP), in conjuction with ISCA '11.

\end{thebibliography}
%\bibliography{paper}

\end{document}