### 2013-09-30 src2/ifl/conclusion.tex

\section{Conclusion} \label{sec:conc}

Obsidian provides a good interface for experimenting with algorithms on GPUs. 
The earlier version described in~\cite{JMT} showed that it is possible to 
generate efficient CUDA code from the kind of high level descriptions we are 
interested in. For the kernel level, the work in progress described in this paper enhances the  expressive power of Obsidian, extending the range of algorithms that can be described, as well as the degree of control exercised by the user. 
Future work will concentrate on improving the performance of the resulting 
applications, as well as on support for the kernel coordination level.