### 2013-09-30 src2/ifl/paper.tex

\documentclass[10pt]{llncs}

\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{subfigure}

\usepackage{alltt}
\usepackage{placeins}

\usepackage{fancyvrb}

   \DefineVerbatimEnvironment
     {code}{Verbatim}
     {fontsize=\small,xleftmargin=0.6cm,samepage=true}

\begin{document}

\title{Obsidian:\\A Domain Specific Embedded Language\\for Parallel Programming of Graphics Processors}

\author{Joel Svensson \and Mary Sheeran \and Koen Claessen}
\institute{Chalmers University of Technology, G\"oteborg, Sweden}

\date{}
\maketitle
\begin{abstract}

We present a domain specific language, embedded in Haskell, for general
purpose parallel programming on GPUs. Our intention is to explore the use of
{\em connection patterns} in parallel programming. We briefly present our
earlier work on hardware generation, and outline the current state of GPU
architectures and programming models. Finally, we present the current status
of the {\em Obsidian} project, which aims to make GPU programming easier,
without relinquishing detailed control of GPU resources. Both a programming
example and some details of the implementation are presented. This is a
report on work in progress.

\end{abstract}
\input{intro}
\input{combinators}   %% uncomment when file available
\input{GPUs}           %% I decided to put a separate section on GPUs, rather
                       %% than trying to cover it in the intro.

\input{obsidian}
\input{results}
\input{related}
\input{futurework}
\input{conclusion}
\bibliographystyle{plain}
\bibliography{ifl08}

\end{document}