### 2013-09-30 src2/ifl/paperThesis.tex

%\documentclass[10pt]{llncs}

%\usepackage{amsmath}
%\usepackage{graphicx}
%\usepackage{subfigure}

%\usepackage{alltt}
%\usepackage{placeins}

%\usepackage{fancyvrb}

   \DefineVerbatimEnvironment
     {code}{Verbatim}
     {fontsize=\small,xleftmargin=0.6cm,samepage=true}

%\begin{document}

%\title{Obsidian:\\A Domain Specific Embedded Language\\for Parallel Programming of Graphics Processors}

%\author{Joel Svensson \and Mary Sheeran \and Koen Claessen}
%\institute{Chalmers University of Technology, G\"oteborg, Sweden}

%\date{}
%\maketitle

\section*{Abstract}

We present a domain specific language, embedded in Haskell, for general
purpose parallel programming on GPUs. Our intention is to explore the use of
{\em connection patterns} in parallel programming. We briefly present our
earlier work on hardware generation, and outline the current state of GPU
architectures and programming models. Finally, we present the current status
of the {\em Obsidian} project, which aims to make GPU programming easier,
without relinquishing detailed control of GPU resources. Both a programming
example and some details of the implementation are presented. This is a
report on work in progress.

%\end{abstract}
\input{./ifl/intro}
\input{./ifl/combinators}   %% uncomment when file available
\input{./ifl/GPUs}           %% I decided to put a separate section on GPUs, rather
                       %% than trying to cover it in the intro.

\input{./ifl/obsidian}
\input{./ifl/results}
\input{./ifl/related}
\input{./ifl/futurework}
\input{./ifl/conclusion}
%\bibliographystyle{plain}
%\bibliography{ifl08}

%\end{document}