### 2013-11-12 blad/page1.tex
update

\thispagestyle{empty}

\begin{center}

\includegraphics{ChalmGUtextsvEng.eps}

\vfill

{\huge Embedded Languages for Data-Parallel Programming}

\vfill

{\LARGE \bf Bo Joel Svensson}

\vfill
The defence of this thesis for the degree of Doctor of Philosophy\linebreak
will be held in room EA, D&IT building,\linebreak
R\"annv\"agen 6B, Chalmers University of Technology,\linebreak
on {\bf December 16, 2013} {\bf at 10.00}.
\linebreak\linebreak
The defence will be held in English.

\vfill

Opponent: Professor Stephen A. Edwards,\linebreak
Department of Computer Science\linebreak
Columbia University\linebreak

\vfill

Department of Computer Science and Engineering\\
Chalmers University of Technology and G\"oteborg University\\
SE-412 96 G\"oteborg, Sweden\\
Telephone: +46 (0)31-772 1000

\vfill

\includegraphics{ChalmGUmarke.eps}

\end{center}

\newpage