### 2013-11-12 blad/page2.tex

\thispagestyle{empty}

\section*{Abstract}

\vspace{5mm}

\begin{small}
Computers today are becoming more and more parallel. General purpose processors (CPUs) 
have multiple processing cores and Single Instruction Multiple Data (SIMD) units for 
data-parallelism. Graphics processors (GPUs) bring massive parallelism at the cost of 
being harder to program than CPUs. This thesis applies embedded language methodology 
to data-parallel programming. Two embedded languages are presented, Obsidian for general 
purpose GPU programming and EmbArBB for data-parallel programming across platforms.

CPUs and GPUs get more parallel resources with each new generation. The question of how to 
efficiently program these processors arises. We are after efficiency both in programmer 
productivity and in application performance. Using embedded languages allows us to experiment
with what abstractions to present to the programmer at relatively little effort.

Obsidian is an embedded language for general purpose programming of GPUs. We try to strike 
a balance between high level, productivity increasing abstractions and low-level 
control needed for performance. The Obsidian programming model mirrors the GPU architecture 
and the programmer is constrained into writing GPU-friendly code. Hierarchy level 
polymorphic library functions are supplied to make these constraints feel less obtrusive. 
Obsidian programs are compiled into CUDA C code. This compilation is based on a simple 
and elegant monad reification technique.

In cases where the programmer is not interested in low-level details 
or wants the program to run over a range of hardware, a higher level language can be used.
EmbArBB is a Haskell embedding or the Intel ArBB system. EmbArBB relies on the ArBB system 
to generate code (via a Just-In-Time compiler) to a range of hardware.

EmbArBB embeds a preexisting library for data-parallelism into Haskell and we obtain 
very good performance at little implementation effort. This performance comes from the 
expertise and effort put into the ArBB system and that we get for free. Embedding ArBB is 
a way to provide these benefits to the Haskell programmer and a way to increase usefulness
of an existing system by opening it up to a wider audience. Obsidian is very different; it 
is not based on a set of high-level parallel primitives. The Obsidian programmer 
can implement these primitives in different ways and then select the best one. 
We have obtained very good performance in case studies involving reductions. Obsidian 
programs are also more terse and composable, compared to CUDA.

\vspace{5mm}

\noindent

\textbf{Keywords:} Data-parallelism, Embedded languages, Functional Programming, Graphics Processing Units

\end{small}
\newpage